---
title: "HNYC Full Run EDA - Version 2"
author: "Gerald"
date: "02/19/2020"
output:
  html_document:
    df_print: paged
---

***

## Script Set-up
Code chunks will be hidden in html output. Look at the rmarkdown file for setting-up.

```{r setup files, include = FALSE}
# Load Files
mn_output <- readRDS("mn_output.rds")
bk_output <- readRDS("bk_output.rds")
```

```{r setup libraries, include = FALSE}
# Load Necessary Libraries
#install.packages("pacman")
library(pacman)
p_load(dplyr,
       DT,
       ggplot2,
       hunspell,
       plotly,
       psych)
```

```{r load GIS libraries, include = FALSE}
# Load libraries for spatial data
p_load(ggmap,
       rgdal,
       rgeos,
       maptools,
       tmap)
```

```{r prepare shapefiles, include = FALSE}
# Load Shapefiles
mn_map <- readOGR(dsn = "mn_shapefiles", layer = "Manhattan_1910")
bk_map <- readOGR(dsn = "bk_shapefiles", layer = "Brooklyn_1910")
```

```{r, include = FALSE}
# Dan's function - Change ED's factor format to numeric
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
```

***

## Table of Contents

* Section 1 - Evaluation of Street Dictionaries & Street Cleaning Function
* Section 2 - EDA of output from the Manhattan cleaning script
* Section 3 - EDA of output from the Brooklyn cleaning Script
* Section 4 - Other Issues
  * Missing EDs in Shapefile and `sample` dataset
  * NA street_add and Best matches
  * Best match to "0"
* Appendix
  * Examples of Different Result Types
  * Manual Matching

***

### Section 1 - Evaluation of Street Dictionaries & Street Cleaning Function

**Street Cleaning Function was editted by Chang on 7th February 2020**  

The following Changes were made:

- The combined Street Dictionary had several issues regarding "STR", "ST" and "STREET". They have now been resolved.

The Street Dictionary (combined and full) for Manhattan and Brooklyn had **several abnormalities**. From manually checking, the issues that were identified include:

- Inconsistent String Pattern
  - e.g. "4 W", "W 4"
  - e.g. "E    97" (Abnormal spacing in between words)
- "ST" seems to have been unintentionally removed from several street names
  - e.g. "CHENUT", "CHENUT ST", "HOUON E", " W HOUON ST", "FIR AVE", "EA END AVE", "CHRIIE", "UYVESANT ST", "CHRISOPHER"
- Possible Misspellings
  - e.g. "WILLET", "WILLETT", "GOERICK", "GOERCK", "EDGECOMBE", "EDGECOMB"
- Inconsistent Street Naming
  - e.g. "LEXINGTON AVE", "LEX AVE"
- Random Street Names
  - e.g. "B", "C", "WE ST", "NYC HOME AGED  AMP; INFIRM", "NY ASYLUM SIERS ST DOMINIC", "MANHATTAN ATE HOSPITAL WARDS ISLAND", "HOME AGED LITTLE SIERS POOR"
- Apostrophe S issues
  - e.g."MARK\\'S PL", "RANDALL\\'S ISLAND PARK", "BEDLOE\\'S ISLAND" (Apostrophe S)
- some NA and NULL values were present

***

### Section 2 - EDA of output from the Manhattan cleaning script

```{r Mn EDA script, include = FALSE}
# Manhattan EDA

mn_output

# How many EDs in Manhattan
mn_output$ED %>% unique() %>% length()

# How many Microfilms in Manhattan
mn_output$microfilm %>% unique() %>% length()

# How many unique Street Names in Manhattan
mn_output$street_add %>% unique() %>% length()

# Street Names
mn_output$best_match %>% unique()

# Total number of entries from sample dataset in Manhattan
nrow(mn_output)

# Tabulate Result Type in Manhattan
# 1 of 6 possible match types: (1) Perfect Match, (2) Identical Match, (3) Singular Mode, (4) Multiple Modes, (5) NA mode, (6) No match. Refer to documentation for more details.
table(mn_output$result_type)

# Tabulate Flagged Filled Streets in Manhattan
table(mn_output$flg_filled_st)

# Tabulate Flagged House Number Cleaned in Manhattan
table(mn_output$flag_hn_cleaned)

# Tabulate Flagged Filled House Numbers in Manhattan
table(mn_output$flg_filled_hn)
```

#### Manhattan's Output

From the sample dataset, an output of 557,357 rows by 33 columns was derived. 

There were 1,478 Enumeration Districts, 1,276 microfilms and 8,557 unique street names within the dataset.

##### Pre-Street-Dictionary Edit

The results from the Street Matching function were as follow:

- 28% Perfect Match
- 12% Identical Match
- 47.5% Singular Mode
- 3% Multiple Modes
- 1.2% NAs
- 0.5% No Match

##### Post-Street-Dictionary Edit (02/07/2020)

The results from the Street Matching function were as follow:

- 34.7% Perfect Match
- 25.4% Identical Match
- 29.8% Singular Mode
- 2% Multiple Modes
- 0.8% NAs
- 0.3% No Match

**Significant improvements were observed** in `result types` 1 and 2 after updating the Street Dictionary

- 5% increase in perfect matches
- 12% increase in identical matches

```{r}
# Distribution of Result Type in Manhattan
plot(table(mn_output$result_type), 
     type = "h", 
     col = c("blue", "red", "orange", "purple", "green", "pink"), 
     lwd = 10,
     main = "Updated Result Type for Manhattan",
     ylab = "Count",
     xlab = "Result Type")
```

##### Problematic Result Types still exist (e.g. 5 and 6 result types)

```{r}
# Get a list of NA and no matches
mn_56_result_list <- mn_output %>% filter(result_type == 5 | result_type == 6) %>% select(township, ED, street_add, best_match, result_type) %>% arrange(ED) %>% dplyr::distinct()
datatable(mn_56_result_list, options = list(autoWidth = TRUE)) %>%
  DT::formatStyle(columns = 1:5, fontSize = '70%')
```

#### Problematic EDs
Out of the problematic Street Matches (result type 5 and type 6), there is a trend of some EDs being more problematic than others, i.e. more entries of 5 or 6 within certain Enumeration Districts.

```{r}
# Problematic EDs for Result Type 5
# Treshold set to 50 (arbitrarily decided)
mn_output %>% filter(result_type == 5) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(19)
```

```{r}
# Problematic EDs for Result Type 6
# Treshold set to 50 (arbitrarily decided)
mn_output %>% filter(result_type == 6) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(8)
```

```{r}
# Average result_type by ED
average_result_type_mn <- mn_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(mean_result_type = mean(result_type))

average_result_type_mn$ED <- as.numeric(average_result_type_mn$ED)

mn_result_type_plot <- ggplot(average_result_type_mn, 
                              aes(x = ED, y = mean_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Average Result Type", title = "Mean Result Type by ED (MN)"
  ) +
  geom_vline(xintercept = 1471, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1652, color = "red", linetype = "dotted")

ggplotly(mn_result_type_plot)
```

```{r}
# Standard Deviation of result_type by ED
sd_result_type_mn <- mn_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(sd_result_type = sd(result_type))

sd_result_type_mn$ED <- as.numeric(sd_result_type_mn$ED)

mn_sd_result_type_plot <- ggplot(sd_result_type_mn, 
                              aes(x = ED, y = sd_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Standard Deviation of Result Type", title = "Standard Deviation of Result Type by ED (MN)"
  ) +
  geom_vline(xintercept = 1394, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1652, color = "red", linetype = "dotted")

ggplotly(mn_sd_result_type_plot)
```

```{r prepare data for MN, include = FALSE}
# Aggregate the mean of result type, i.e. street matching success rate
# The lower the better
mn_output_result_type <- aggregate(result_type ~ ED, FUN = mean, data = mn_output)
mn_output_result_type$ED <- as.numeric(mn_output_result_type$ED)

# Change Manhattan Shapefile ED variable to numeric format (Dan's function)
mn_map@data$ED <- as.numeric.factor(mn_map@data$ED)

# Left join the 2 datasets
mn_map@data <- left_join(mn_map@data, mn_output_result_type, by = c('ED' = 'ED'))
```

```{r}
# Map for Result Type in Manhattan
tmap_mode("view")
tm_shape(mn_map) + tm_polygons("result_type")
```

#### Fill Down Function - Manhattan

Out of the 28% of Perfect Matches (156k entries) 1,717 entries were matched via the Fill Down function `03_Matched_Street_Fill_Down` - 1.1%.

#### Flagged House Numbers - Manhattan

Out of 557,357 entries, 159,669 entries were flagged for house number changes (i.e. 0 or 1) - 28.6%.

Previously, 7,350 entries were flagged with "1" meaning the house number was editted by the function. That has now grown to 7,520 (post street cleaning function edits).

**This flag refers to:** The initial household number might have been "195-7", "34 TO 36" or "112 114". There would be a split between the house numbers and `hn_1` will be 195 and `hn_2` will be 197 as in the first example.

```{r}
# Problematic Enumeration Districts for House Number Cleaning
mn_output %>% filter(flag_hn_cleaned == 1) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(24)
```

Out of all the house hold number entries, 364,730 entries were filled in via the Fill Down function `05_House_Number_Fill_Down` - 65.4%.

That has increased to 368,948 - 66.1% after the street cleaning function was editted.

***

### Section 3 - EDA of output from the Brooklyn cleaning Script

```{r Bk EDA script, include = FALSE}
# Brooklyn EDA

bk_output

# How many EDs in Brooklyn
bk_output$ED %>% unique() %>% length()

# How many Microfilms in Brooklyn
bk_output$microfilm %>% unique() %>% length()

# How many unique Street Names in Brooklyn
bk_output$street_add %>% unique() %>% length()

# Street Names Brooklyn
bk_output$best_match %>% unique() %>% head(200)

# Total number of entries from sample dataset in Brooklyn
nrow(bk_output)

# Tabulate Result Type in Brooklyn
# 1 of 6 possible match types: (1) Perfect Match, (2) Identical Match, (3) Singular Mode, (4) Multiple Modes, (5) NA mode, (6) No match. Refer to documentation for more details.
table(bk_output$result_type)

# Tabulate Flagged Filled Streets in Brooklyn
table(bk_output$flg_filled_st)

# Tabulate Flagged House Number Cleaned in Brooklyn
table(bk_output$flag_hn_cleaned)

# Tabulate Flagged Filled House Numbers in Brooklyn
table(bk_output$flg_filled_hn)
```

#### Brooklyn Output

From the sample dataset, an output of 371,833 rows by 33 columns was derived. 

There were 1,106 Enumeration Districts, 1,527 microfilms and 7,076 unique street names within the dataset.

##### Pre-Street-Dictionary Update

The results from the Street Matching function are as follow:

- 76.5% Perfect Match
- 7.9% Identical Match
- 9.5% Singular Mode
- 1.5% Multiple Modes
- 1.3% NAs
- 0.8% No Match

##### Post-Street-Dictionary Update

- 78.7% Perfect Match
- 8.3% Identical Match
- 7.3% Singular Mode
- 1.5% Multiple Modes
- 1.0% NAs
- 0.7% No Match

In contrast to Manhattan, there is a larger proportion of Perfect Matches.

```{r}
# Distribution of Result type in Brooklyn
plot(table(bk_output$result_type), 
     type = "h", 
     col = c("blue", "red", "orange", "purple", "green", "pink"), 
     lwd = 10,
     main = "Result Type for Brooklyn",
     ylab = "Count",
     xlab = "Result Type")
```

```{r}
# Get a list of NA and no matches
bk_56_result_list <- bk_output %>% filter(result_type == 5 | result_type == 6) %>% select(township, ED, street_add, best_match, result_type) %>% arrange(ED) %>% dplyr::distinct()
bk_56_result_list
```

#### Problematic EDs - Brooklyn

Out of the problematic Street Matches (result type 5 and type 6), there is a trend of some EDs being more problematic than others, i.e. more entries of 5 or 6 within certain Enumeration Districts.

```{r}
# Problematic EDs for Result Type 5
# Treshold set to 50 (arbitrarily decided)
bk_output %>% filter(result_type == 5) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(18)
```

```{r}
# Problematic EDs for Result Type 6
# Treshold set to 50 (arbitrarily decided)
bk_output %>% filter(result_type == 6) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(10)
```

```{r}
# Average result_type by ED
average_result_type_bk <- bk_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(mean_result_type = mean(result_type))

average_result_type_bk$ED <- as.numeric(average_result_type_bk$ED)

bk_result_type_plot <- ggplot(average_result_type_bk, 
                              aes(x = ED, y = mean_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Average Result Type", title = "Mean Result Type by ED (BK)"
  ) +
  geom_vline(xintercept = 1103, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1406, color = "red", linetype = "dotted")

ggplotly(bk_result_type_plot)
```

```{r}
# Standard Deviation of result_type by ED
sd_result_type_bk <- bk_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(sd_result_type = sd(result_type))

sd_result_type_bk$ED <- as.numeric(sd_result_type_bk$ED)

bk_sd_result_type_plot <- ggplot(sd_result_type_bk, 
                              aes(x = ED, y = sd_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Standard Deviation of Result Type", title = "Standard Deviation of Result Type by ED (BK)"
  ) +
  geom_vline(xintercept = 1103, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1406, color = "red", linetype = "dotted")

ggplotly(bk_sd_result_type_plot)
```

```{r prepare data for BK, include = FALSE}
# Aggregate the mean of result type, i.e. street matching success rate
# The lower the better
bk_output_result_type <- aggregate(result_type ~ ED, FUN = mean, data = bk_output)
bk_output_result_type$ED <- as.numeric(bk_output_result_type$ED)

# Change Brooklyn Shapefile ED variable to numeric format
bk_map@data$ED <- as.numeric.factor(bk_map@data$ED)

# Left join the 2 datasets
bk_map@data <- left_join(bk_map@data, bk_output_result_type, by = c('ED' = 'ED'))
```

```{r}
# Map for Result Type in Brooklyn
tmap_mode("view")
tm_shape(bk_map) + tm_polygons("result_type")
```

#### Fill Down Function - Brooklyn

Out of the 76.5% of Perfect Matches (284k entries) 1,005 entries were matched via the Fill Down function `03_Matched_Street_Fill_Down` - A much smaller percentage than Manhattan. This suggests that the recording of entries might be more accurate in Brooklynn or that the street directory is more well-developed. Nevertheless, errors are still present.

#### Flagged House Numbers - Brooklyn

Out of 371,833 entries, 179,303 entries were flagged for house number changes (i.e. 0 or 1) - 48.2%.

2,563 entries were flagged with "1" meaning the house number was editted by the function. Brooklyn also has several abnormal initial house numbers. E.g. "222 1/2", "192 3TH". Otherwise, the issues are similar to Manhattan.

```{r}
# Problematic Enumeration Districts for House Number Cleaning
bk_output %>% filter(flag_hn_cleaned == 1) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(11)
```

Out of all the house hold number entries, 182,950 entries were filled in via the Fill Down function `05_House_Number_Fill_Down` - 49.1%.

***

## Section 4 - Other Issues

#### There are missing EDs in the Dataset
    - There are missing EDs in the `sample` dataset and missing EDs in the `shapefiles`
    - The EDs could be names wrongly
        - In the Shapefile, the max ED is 1480 but the max ED in the output runs till 1941

```{r}
# Missing EDs in Manhattan

# Manhattan ED list
mn_ed_list <- mn_map@data$ED %>% sort()

mn_output_ed <- mn_output$ED %>% unique() %>% as.numeric() %>% sort()
```

```{r}
# Which EDs are in the Shapefile but missing from the Manhattan dataset
mn_ed_list[!(mn_output_ed %in% mn_ed_list)]
length(mn_ed_list[!(mn_output_ed %in% mn_ed_list)])
```

```{r}
# Which EDs are in the dataset but missing from the Manhattan shapefile
mn_output_ed[!(mn_ed_list %in% mn_output_ed)]
length(mn_output_ed[!(mn_ed_list %in% mn_output_ed)])
```

```{r}
# Missing EDs in Brooklyn

# Brooklyn ED list
bk_ed_list <- bk_map@data$ED %>% sort()

# Brooklyn Output EDs
bk_output_ed <- bk_output$ED %>% unique() %>% as.numeric() %>% sort()
```

```{r}
# Which EDs are in the Shapefile but missing from the Brooklyn dataset
bk_ed_list[!(bk_output_ed %in% bk_ed_list)]
length(bk_ed_list[!(bk_output_ed %in% bk_ed_list)])
```

```{r}
# Which EDs are in the dataset but missing from the Brooklyn shapefile
bk_output_ed[!(bk_ed_list %in% bk_output_ed)]
length(bk_output_ed[!(bk_ed_list %in% bk_output_ed)])
```

<br/>

#### Problematic Matches - Manhattan

There are 39,311 entries with NA `best_match`. Their `street_add` entries are also NA. There are 795 EDs with these problems.

* The original `sample` data might be corrupted. Or have missing street addresses

```{r}
mn_output %>% filter(is.na(best_match)) %>% sample_n(5)
```

```{r}
na_mn_output <- mn_output %>% filter(is.na(best_match)) %>% select(ED) %>% unique()

na_mn_output <- na_mn_output %>% mutate('NA best match' = 1)

datatable(na_mn_output, options = list(autoWidth = TRUE)) %>%
  DT::formatStyle(columns = 1:5, fontSize = '70%')
```

```{r}
na_mn_map <- merge(mn_map, na_mn_output, 'ED', 'ED')
```

```{r}
# Highlight Example
example <- mn_output %>% filter(ED == '0227')
example[41:60,]
```

```{r}
# Map for NA Street Address / Best Match in Manhattan
tmap_mode("view")
tm_shape(na_mn_map) + tm_polygons("NA best match")
```

* Street Dictionary has "0", which is deriving this issue.
* There are 80 such streets affected

```{r}
# Another problematic best match is "0"
mn_output %>% filter(best_match == 0) %>% sample_n(10)
```

<br/>

#### Problematic Matches - Brooklyn

There are 9,264 entries with NA `best_match`. Their `street_add` entries are also NA. There are 427 EDs with these problems.

```{r}
bk_output %>% filter(is.na(best_match)) %>% sample_n(5)
```

```{r}
na_bk_output <- bk_output %>% filter(is.na(best_match)) %>% select(ED) %>% unique()

na_bk_output <- na_bk_output %>% mutate('NA best match' = 1)

datatable(na_bk_output, options = list(autoWidth = TRUE)) %>%
  DT::formatStyle(columns = 1:5, fontSize = '70%')
```

```{r}
na_bk_map <- merge(bk_map, na_bk_output, 'ED', 'ED')
```

```{r}
# Map for NA Street Address / Best Match in Brooklyn
tmap_mode("view")
tm_shape(na_bk_map) + tm_polygons("NA best match")
```

* Brooklyn has fewer of such "0" problems.
* 11 unique streets are affected

```{r}
# Another problematic best match is "0"
bk_output %>% filter(best_match == 0) %>% sample_n(10)
```

***

## Appendix

#### Examples of the different result types

##### Perfect Match
```{r}
# Sample of Result Type 1 (Perfect Match)
mn_output %>% filter(result_type == 1) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### Identical Match
```{r}
# Sample of Result Type 2 (Identical Match)
mn_output %>% filter(result_type == 2) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### Singular Mode Match
```{r}
# Sample of Result Type 3 (Singular Mode)
mn_output %>% filter(result_type == 3) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### Multiple Modes Match
```{r}
# Sample of Result Type 4 (Multiple Modes)
mn_output %>% filter(result_type == 4) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### "NA" Match
```{r}
# Sample of Result Type 1 (NA)
mn_output %>% filter(result_type == 5) %>% select(ED, street_add, best_match, result_type) %>% head(3)
```

##### No Match
```{r}
# Sample of Result Type 6 (no match)
mn_output %>% filter(result_type == 6) %>% select(ED, street_add, best_match, result_type) %>% head(4)
```

#### Appendix - Trouble Shooting / Manual Checking

List of current issues:
* There are "NA" `result_type`, which are not captured in our 5 and 6s
  * These entries typically have missing `street_add`
  * 38,237 such entries in Manhattan (6.86%)
  * 8,462 such entries in Brooklyn (2.27%)
  
* There are entries that match to "0" in the street dictionary
  * These entries typically have `result_type` 4
  * 1,018 such entries in Manhattan (0.18%)
  * 1,032 such entries in Brooklyn (0.27%)
  
* 5 and 6 matches

  * String matches reliably close but still produces a 6 output
    * EAST HOUSTON STREET --> E HOUSTON (`result_type` == 6)
    * CLINTON --> CLINTON AVE
    * BROOM STREET --> BROOME
    * DEKALB & AVENUE --> DE KALB AVE
    * Recommendation: Confident manual match
    
  * String matches part of the best match but still produces a 5 output
    * EAST 116 ST BEBRA FIRST AND PLED --> E 116
    * Recommendation: Confident manual match
    
  * Numerical String Unreliable match (produces either 5 or 6 `result_type`)
    * East 110 --> E 116
    * Recommendation: Investigate before manual match
    
  * Spelt out form numerical matches
    * EAST ONE HUNDRED & TWENTIETH ST --> E 119
    * EAST ONE HUNDRED AND THIRTEENTH --> E 113
    * EAST 99THTREET --> E 99
    * Recommendation: Update street dictionary to turn spelt out streets into numerical format. (Current function should already be doing this, especially the EAST to E). Confident manual match.
    
  * Completely random output with either 5 or 6 `result_type`
    * ?? --> E120
    * 444 --> 1 AVE
    * RIRIE TWNRD 117 ST 118 ST --> PLEASANT AVE
    * Recommendation: NULL

```{r, include = FALSE}
# Create a list for manual checking
combine_56_result_list <- rbind(mn_56_result_list, bk_56_result_list)
#write.csv(combine_56_result_list, "manual_check.csv")
```

```{r}
combine_56_result_list
```

* Is it possible to use Hunspell's spell check function to reduce the number of poor best matches?
  * There are 771 words that were returned from the hunspell function
  * There is a strong use case for certain streets
    * GREENRICH --> GREENWICH
    * CARRAL --> CANAL
    * COYAL --> CANAL
    * CHRYSTEE --> CHRYSTIE
    * ELDREDGE --> ELDRIDGE
    * NAPOLK ST --> NORFOLK
    * TEX --> LEX AVE

```{r}
# Use hunspell to get a list of possible mispelt streets
misspelt <- hunspell_parse(combine_56_result_list$street_add) %>% unlist() %>% hunspell() %>% unlist()
misspelt_list <- misspelt[misspelt != "character(0)"]
misspelt_list %>% head(100)
```