---
title: "HNYC Full Run EDA - Version 2"
author: "Gerald"
date: "02/07/2020"
output:
  html_document:
    df_print: paged
---

***

### EDA for Street Cleaning Function & Strict Dictionary

**Street Cleaning Function editted by Chang**  
Editted on 7th Feburary 2020

Following Changes were made:

The combined Street Dictionary had several issues regarding "STR", "ST" and "STREET". They have been resolved.

The Street Dictionary (combined and full) for Manhattan and Brooklyn had **several abnormalities**. From manual checking the issues that were identified included:

- Inconsistent String Pattern
  - e.g. "4 W", "W 4", "E    97" (Abnormal spacing)
- "ST" seems to have been unintentionally removed from several street names
  - e.g. "CHENUT", "CHENUT ST", "HOUON E", " W HOUON ST", "FIR AVE", "EA END AVE", "CHRIIE", "UYVESANT ST", "CHRISOPHER"
- Possible Misspellings
  - e.g. "WILLET", "WILLETT", "GOERICK", "GOERCK", "EDGECOMBE", "EDGECOMB"
- Inconsistent Street Naming
  - e.g. "LEXINGTON AVE", "LEX AVE"
- Random Street Names
  - e.g. "B", "C", "WE ST", "NYC HOME AGED  AMP; INFIRM", "NY ASYLUM SIERS ST DOMINIC", "MANHATTAN ATE HOSPITAL WARDS ISLAND", "HOME AGED LITTLE SIERS POOR"
- Apostrophe S issues
  - e.g."MARK\\'S PL", "RANDALL\\'S ISLAND PARK", "BEDLOE\\'S ISLAND" (Apostrophe S)

The following issues have not been resolved:

- NA or NULL

***

### EDA of output from the Manhattan cleaning script

```{r setup files, include = FALSE}
# Load Files
mn_output <- readRDS("mn_output.rds")
bk_output <- readRDS("bk_output.rds")
```

```{r setup libraries, include = FALSE}
# Load Necessary Libraries
#install.packages("pacman")
library(pacman)
p_load(dplyr,
       ggplot2,
       plotly,
       psych)
```

```{r Mn EDA script, include = FALSE}
# Manhattan EDA

mn_output

# How many EDs in Manhattan
mn_output$ED %>% unique() %>% length()

# How many Microfilms in Manhattan
mn_output$microfilm %>% unique() %>% length()

# How many unique Street Names in Manhattan
mn_output$street_add %>% unique() %>% length()

# Street Names
mn_output$best_match %>% unique()

# Total number of entries from sample dataset in Manhattan
nrow(mn_output)

# Tabulate Result Type in Manhattan
# 1 of 6 possible match types: (1) Perfect Match, (2) Identical Match, (3) Singular Mode, (4) Multiple Modes, (5) NA mode, (6) No match. Refer to documentation for more details.
table(mn_output$result_type)

# Tabulate Flagged Filled Streets in Manhattan
table(mn_output$flg_filled_st)

# Tabulate Flagged House Number Cleaned in Manhattan
table(mn_output$flag_hn_cleaned)

# Tabulate Flagged Filled House Numbers in Manhattan
table(mn_output$flg_filled_hn)
```

#### Manhattan

From the sample dataset, an output of 557,357 rows by 33 columns was derived. 

There were 1,478 Enumeration Districts, 1,276 microfilms and 8,557 unique street names within the dataset.

##### Pre-Street-Cleaning Function Edit

The results from the Street Matching function were as follow:

- 28% Perfect Match
- 12% Identical Match
- 47.5% Singular Mode
- 3% Multiple Modes
- 1.2% NAs
- 0.5% No Match

##### Post-Street-Cleaning Function Edit

The results from the Street Matching function were as follow:

- 34.7% Perfect Match
- 25.4% Identical Match
- 29.8% Singular Mode
- 2% Multiple Modes
- 0.8% NAs
- 0.3% No Match

**Significant improvements observed** in `result types` 1 and 2 after improving the Street Cleaning Function.

- 5% increase in perfect matches
- 12% increase in identical matches

```{r}
# Distribution of Result Type in Manhattan
plot(table(mn_output$result_type), 
     type = "h", 
     col = c("blue", "red", "orange", "purple", "green", "pink"), 
     lwd = 10,
     main = "Result Type for Manhattan",
     ylab = "Count",
     xlab = "Result Type")
```

#### For Reference, these are examples of the different result types:

##### Perfect Match:
```{r}
# Sample of Result Type 1 (Perfect Match)
mn_output %>% filter(result_type == 1) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### Identical Match:
```{r}
# Sample of Result Type 2 (Identical Match)
mn_output %>% filter(result_type == 2) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### Singular Mode Match:
```{r}
# Sample of Result Type 3 (Singular Mode)
mn_output %>% filter(result_type == 3) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### Multiple Modes Match:
```{r}
# Sample of Result Type 4 (Multiple Modes)
mn_output %>% filter(result_type == 4) %>% select(ED, street_add, best_match, result_type) %>% head(1)
```

##### "NA" Match:
```{r}
# Sample of Result Type 1 (NA)
mn_output %>% filter(result_type == 5) %>% select(ED, street_add, best_match, result_type) %>% head(3)
```

##### No Match:
```{r}
# Sample of Result Type 6 (no match)
mn_output %>% filter(result_type == 6) %>% select(ED, street_add, best_match, result_type) %>% head(4)
```

#### Problematic EDs
Out of the problematic Street Matches (result type 5 and type 6), there is a trend of some EDs being more problematic than others, i.e. more entries of 5 or 6 within certain Enumeration Districts.

```{r}
# Problematic EDs for Result Type 5
# Treshold set to 50 (arbitrarily decided)
mn_output %>% filter(result_type == 5) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(19)
```

```{r}
# Problematic EDs for Result Type 6
# Treshold set to 50 (arbitrarily decided)
mn_output %>% filter(result_type == 6) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(8)
```

```{r}
# Average result_type by ED
average_result_type_mn <- mn_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(mean_result_type = mean(result_type))

average_result_type_mn$ED <- as.numeric(average_result_type_mn$ED)

mn_result_type_plot <- ggplot(average_result_type_mn, 
                              aes(x = ED, y = mean_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Average Result Type", title = "Mean Result Type by ED (MN)"
  ) +
  geom_vline(xintercept = 1471, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1652, color = "red", linetype = "dotted")

ggplotly(mn_result_type_plot)
```

```{r}
# Standard Deviation of result_type by ED
sd_result_type_mn <- mn_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(sd_result_type = sd(result_type))

sd_result_type_mn$ED <- as.numeric(sd_result_type_mn$ED)

mn_sd_result_type_plot <- ggplot(sd_result_type_mn, 
                              aes(x = ED, y = sd_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Standard Deviation of Result Type", title = "Standard Deviation of Result Type by ED (MN)"
  ) +
  geom_vline(xintercept = 1394, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1652, color = "red", linetype = "dotted")

ggplotly(mn_sd_result_type_plot)
```

#### Fill Down Function - Manhattan

Out of the 28% of Perfect Matches (156k entries) 1,717 entries were matched via the Fill Down function `03_Matched_Street_Fill_Down` - 1.1%.

#### Flagged House Numbers - Manhattan

Out of 557,357 entries, 159,669 entries were flagged for house number changes (i.e. 0 or 1) - 28.6%.

Previously, 7,350 entries were flagged with "1" meaning the house number was editted by the function. That has now grown to 7,520 (post street cleaning function edits).

**This flag refers to:** The initial household number might have been "195-7", "34 TO 36" or "112 114". There would be a split between the house numbers and `hn_1` will be 195 and `hn_2` will be 197 as in the first example.

```{r}
# Problematic Enumeration Districts for House Number Cleaning
mn_output %>% filter(flag_hn_cleaned == 1) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(24)
```

Out of all the house hold number entries, 364,730 entries were filled in via the Fill Down function `05_House_Number_Fill_Down` - 65.4%.

That has increased to 368,948 - 66.1% after the street cleaning function was editted.

***

```{r Bk EDA script, include = FALSE}
# Brooklyn EDA

bk_output

# How many EDs in Brooklyn
bk_output$ED %>% unique() %>% length()

# How many Microfilms in Brooklyn
bk_output$microfilm %>% unique() %>% length()

# How many unique Street Names in Brooklyn
bk_output$street_add %>% unique() %>% length()

# Street Names Brooklyn
bk_output$best_match %>% unique() %>% head(200)

# Total number of entries from sample dataset in Brooklyn
nrow(bk_output)

# Tabulate Result Type in Brooklyn
# 1 of 6 possible match types: (1) Perfect Match, (2) Identical Match, (3) Singular Mode, (4) Multiple Modes, (5) NA mode, (6) No match. Refer to documentation for more details.
table(bk_output$result_type)

# Tabulate Flagged Filled Streets in Brooklyn
table(bk_output$flg_filled_st)

# Tabulate Flagged House Number Cleaned in Brooklyn
table(bk_output$flag_hn_cleaned)

# Tabulate Flagged Filled House Numbers in Brooklyn
table(bk_output$flg_filled_hn)
```

#### Brooklyn

From the sample dataset, an output of 371,833 rows by 33 columns was derived. 

There were 1,106 Enumeration Districts, 1,527 microfilms and 7,076 unique street names within the dataset.

##### Pre-Street-Cleaning Function Edit

The results from the Street Matching function are as follow:

- 76.5% Perfect Match
- 7.9% Identical Match
- 9.5% Singular Mode
- 1.5% Multiple Modes
- 1.3% NAs
- 0.8% No Match

##### Post-Street-Cleaning Function Edit

- 78.7% Perfect Match
- 8.3% Identical Match
- 7.3% Singular Mode
- 1.5% Multiple Modes
- 1.0% NAs
- 0.7% No Match

In contrast to Manhattan, there is a larger proportion of Perfect Matches.

```{r}
# Distribution of Result type in Brooklyn
plot(table(bk_output$result_type), 
     type = "h", 
     col = c("blue", "red", "orange", "purple", "green", "pink"), 
     lwd = 10,
     main = "Result Type for Brooklyn",
     ylab = "Count",
     xlab = "Result Type")
```

#### Problematic EDs - Brooklyn

Out of the problematic Street Matches (result type 5 and type 6), there is a trend of some EDs being more problematic than others, i.e. more entries of 5 or 6 within certain Enumeration Districts.

```{r}
# Problematic EDs for Result Type 5
# Treshold set to 50 (arbitrarily decided)
bk_output %>% filter(result_type == 5) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(18)
```

```{r}
# Problematic EDs for Result Type 6
# Treshold set to 50 (arbitrarily decided)
bk_output %>% filter(result_type == 6) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(10)
```

```{r}
# Average result_type by ED
average_result_type_bk <- bk_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(mean_result_type = mean(result_type))

average_result_type_bk$ED <- as.numeric(average_result_type_bk$ED)

bk_result_type_plot <- ggplot(average_result_type_bk, 
                              aes(x = ED, y = mean_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Average Result Type", title = "Mean Result Type by ED (BK)"
  ) +
  geom_vline(xintercept = 1103, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1406, color = "red", linetype = "dotted")

ggplotly(bk_result_type_plot)
```

```{r}
# Standard Deviation of result_type by ED
sd_result_type_bk <- bk_output %>% 
  select(ED, result_type) %>%
  group_by(ED) %>%
  summarise(sd_result_type = sd(result_type))

sd_result_type_bk$ED <- as.numeric(sd_result_type_bk$ED)

bk_sd_result_type_plot <- ggplot(sd_result_type_bk, 
                              aes(x = ED, y = sd_result_type)) +
  theme_classic() +
  geom_point() + 
  labs(x = "ED", y = "Standard Deviation of Result Type", title = "Standard Deviation of Result Type by ED (BK)"
  ) +
  geom_vline(xintercept = 1103, color = "red", linetype = "dotted") +
  geom_vline(xintercept = 1406, color = "red", linetype = "dotted")

ggplotly(bk_sd_result_type_plot)
```

#### Fill Down Function - Brooklyn

Out of the 76.5% of Perfect Matches (284k entries) 1,005 entries were matched via the Fill Down function `03_Matched_Street_Fill_Down` - A much smaller percentage than Manhattan. This suggests that the recording of entries might be more accurate in Brooklynn or that the street directory is more well-developed. Nevertheless, errors are still present.

#### Flagged House Numbers - Brooklyn

Out of 371,833 entries, 179,303 entries were flagged for house number changes (i.e. 0 or 1) - 48.2%.

2,563 entries were flagged with "1" meaning the house number was editted by the function. Brooklyn also has several abnormal initial house numbers. E.g. "222 1/2", "192 3TH". Otherwise, the issues are similar to Manhattan.

```{r}
# Problematic Enumeration Districts for House Number Cleaning
bk_output %>% filter(flag_hn_cleaned == 1) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(11)
```

Out of all the house hold number entries, 182,950 entries were filled in via the Fill Down function `05_House_Number_Fill_Down` - 49.1%.

***

#### Visualizing the Outputs spatially

```{r load GIS libraries, include = FALSE}
p_load(ggmap,
       rgdal,
       rgeos,
       maptools,
       tmap)
```

```{r prepare shapefiles, include = FALSE}
mn_map <- readOGR(dsn = "mn_shapefiles", layer = "Manhattan_1910")
bk_map <- readOGR(dsn = "bk_shapefiles", layer = "Brooklyn_1910")
```

```{r}
# Dan's function - Change ED's factor format to numeric
as.numeric.factor <- function(x) {as.numeric(levels(x))[x]}
```

```{r prepare data for MN}
# Aggregate the mean of result type, i.e. street matching success rate
# The lower the better
mn_output_result_type <- aggregate(result_type ~ ED, FUN = mean, data = mn_output)
mn_output_result_type$ED <- as.numeric(mn_output_result_type$ED)

# Change Manhattan Shapefile ED variable to numeric format (Dan's function)
mn_map@data$ED <- as.numeric.factor(mn_map@data$ED)

# Left join the 2 datasets
mn_map@data <- left_join(mn_map@data, mn_output_result_type, by = c('ED' = 'ED'))
```

```{r}
# Map for Result Type in Manhattan
tmap_mode("view")
tm_shape(mn_map) + tm_polygons("result_type")
```

```{r prepare data for BK}
# Aggregate the mean of result type, i.e. street matching success rate
# The lower the better
bk_output_result_type <- aggregate(result_type ~ ED, FUN = mean, data = bk_output)
bk_output_result_type$ED <- as.numeric(bk_output_result_type$ED)

# Change Brooklyn Shapefile ED variable to numeric format
bk_map@data$ED <- as.numeric.factor(bk_map@data$ED)

# Left join the 2 datasets
bk_map@data <- left_join(bk_map@data, bk_output_result_type, by = c('ED' = 'ED'))
```

```{r}
# Map for Result Type in Brooklyn
tmap_mode("view")
tm_shape(bk_map) + tm_polygons("result_type")
```

```{r}
# Highlight 2 examples in Brooklyn with poor street name matches
bk_output %>% filter(ED == 1034 | ED == 1039)
```

***

#### Other Problems to Look into

- Missing EDs in the entire Dataset
    - There are missing EDs in the entire dataset
    - The EDs could be names wrongly
        - In the Shapefile, the max ED is 1480 but the max ED in the output runs till 1941

```{r}
# Missing EDs in Manhattan

# Manhattan ED list
mn_ed_list <- mn_map@data$ED %>% sort()

mn_output_ed <- mn_output$ED %>% unique() %>% as.numeric() %>% sort()
```

```{r}
# Which EDs are in the Shapefile but missing from the Manhattan dataset
mn_ed_list[!(mn_output_ed %in% mn_ed_list)]
length(mn_ed_list[!(mn_output_ed %in% mn_ed_list)])
```

```{r}
# Which EDs are in the dataset but missing from the Manhattan shapefile
mn_output_ed[!(mn_ed_list %in% mn_output_ed)]
length(mn_output_ed[!(mn_ed_list %in% mn_output_ed)])
```

```{r}
# Missing EDs in Brooklyn

# Brooklyn ED list
bk_ed_list <- bk_map@data$ED %>% sort()

# Brooklyn Output EDs
bk_output_ed <- bk_output$ED %>% unique() %>% as.numeric() %>% sort()
```

```{r}
# Which EDs are in the Shapefile but missing from the Brooklyn dataset
bk_ed_list[!(bk_output_ed %in% bk_ed_list)]
length(bk_ed_list[!(bk_output_ed %in% bk_ed_list)])
```

```{r}
# Which EDs are in the dataset but missing from the Brooklyn shapefile
bk_output_ed[!(bk_ed_list %in% bk_output_ed)]
length(bk_output_ed[!(bk_ed_list %in% bk_output_ed)])
```

#### Problematic Best Matches - Manhattan

There are 38,174 entries with NA `best_match`. Their `street_add` entries are also NA. There are 792 EDs with these problems.

```{r}
na_mn_output <- mn_output %>% filter(is.na(best_match)) %>% select(ED) %>% unique()

na_mn_output <- na_mn_output %>% mutate('NA best match' = 1)
```

```{r}
na_mn_map <- merge(mn_map, na_mn_output, 'ED', 'ED')
```

```{r}
# Map for NA Street Address / Best Match in Manhattan
tmap_mode("view")
tm_shape(na_mn_map) + tm_polygons("NA best match")
```

```{r}
# Another problematic best match is "0"
mn_output %>% filter(best_match == 0) %>% select(street_add) %>% unique()
```

#### Problematic Best Matches - Brooklyn

There are 9,264 entries with NA `best_match`. Their `street_add` entries are also NA. There are 427 EDs with these problems.

```{r}
na_bk_output <- bk_output %>% filter(is.na(best_match)) %>% select(ED) %>% unique()

na_bk_output <- na_bk_output %>% mutate('NA best match' = 1)
```

```{r}
na_bk_map <- merge(bk_map, na_bk_output, 'ED', 'ED')
```

```{r}
# Map for NA Street Address / Best Match in Brooklyn
tmap_mode("view")
tm_shape(na_bk_map) + tm_polygons("NA best match")
```

```{r}
# Another problematic best match is "0"
bk_output %>% filter(best_match == 0) %>% select(street_add) %>% unique()
```
