---
title: "HNYC Full Run EDA"
author: "Gerald"
output:
  html_document:
    df_print: paged
---

***

### EDA for Street Cleaning Function & Strict Dictionary

**Follow Up Action Required: Check if this is appropriate**  

The Street Cleaning Function has 3 lines of codes that were set to be ignored. 

The combined Street Dictionary has several issues regarding "STR", "ST" and "STREET". This might be a possible cause of those issues. 

```{r}
# Lines of codes that had "##" in the Street Cleaning Function
##x<-gsub("\\bSTR\\b","", x)
##x<-gsub("\\<ST\\>","",x)
##x<-gsub("\\<STREET\\>","",x)
```

The Street Dictionary (combined and full) for Manhattan have **several abnormalities**. From manual checking the issues that were identified include:

- NA
- Inconsistent String Pattern
  - e.g. "4 W", "W 4", "E    97" (Abnormal spacing)
- "ST" seems to have been unintentionally removed from several street names
  - e.g. "CHENUT", "CHENUT ST", "HOUON E", " W HOUON ST", "FIR AVE", "EA END AVE", "CHRIIE", "UYVESANT ST", "CHRISOPHER"
- Possible Misspellings
  - e.g. "WILLET", "WILLETT", "GOERICK", "GOERCK", "EDGECOMBE", "EDGECOMB"
- Inconsistent Street Naming
  - e.g. "LEXINGTON AVE", "LEX AVE"
- Random Street Names
  - e.g. "B", "C", "WE ST", "NYC HOME AGED  AMP; INFIRM", "NY ASYLUM SIERS ST DOMINIC", "MANHATTAN ATE HOSPITAL WARDS ISLAND", "HOME AGED LITTLE SIERS POOR"
- Apostrophe S issues
  - e.g."MARK\\'S PL", "RANDALL\\'S ISLAND PARK", "BEDLOE\\'S ISLAND"

***

### EDA of output from Manhattan cleaning script

```{r setup files, include = FALSE}
# Load Files
mn_output <- readRDS("mn_output.rds")
bk_output <- readRDS("bk_output.rds")
```

```{r setup libraries, include = FALSE}
# Load Necessary Libraries
#install.packages("pacman")
library(pacman)
p_load(dplyr,
       ggplot2,
       psych)
```

```{r Mn EDA script, include = FALSE}
# Manhattan EDA

mn_output

# How many EDs in Manhattan
mn_output$ED %>% unique() %>% length()

# How many Microfilms in Manhattan
mn_output$microfilm %>% unique() %>% length()

# How many unique Street Names in Manhattan
mn_output$street_add %>% unique() %>% length()

# Street Names
mn_output$best_match %>% unique()

# Total number of entries from sample dataset in Manhattan
nrow(mn_output)

# Tabulate Result Type in Manhattan
# 1 of 6 possible match types: (1) Perfect Match, (2) Identical Match, (3) Singular Mode, (4) Multiple Modes, (5) NA mode, (6) No match. Refer to documentation for more details.
table(mn_output$result_type)

# Tabulate Flagged Filled Streets in Manhattan
table(mn_output$flg_filled_st)

# Tabulate Flagged House Number Cleaned in Manhattan
table(mn_output$flag_hn_cleaned)

# Tabulate Flagged Filled House Numbers in Manhattan
table(mn_output$flg_filled_hn)
```

#### Manhattan

From the sample dataset, an output of 557,357 rows by 33 columns was derived. There are 1,478 Enumeration Districts, 1,276 microfilms and 8,557 unique street names within the dataset.

The results from the Street Matching function are as follow:

- 28% Perfect Match
- 12% Identical Match
- 47.5% Singular Mode
- 3% Multiple Modes
- 1.2% NAs
- 0.5% No Match

Out of the 28% of Perfect Matches (156k entries) 1,717 entries were matched via the Fill Down function `03_Matched_Street_Fill_Down` - 1.1%.

Out of the problematic Street Matches (result type 5 and type 6), there is a trend of some EDs being more problematic than others, i.e. more entries of 5 or 6 within certain Enumeration Districts.

```{r}
# Problematic EDs for Result Type 5
# Treshold set to 50 (arbitrarily decided)
mn_output %>% filter(result_type == 5) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(29)
```

```{r}
# Problematic EDs for Result Type 6
# Treshold set to 50 (arbitrarily decided)
mn_output %>% filter(result_type == 6) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(6)
```

Out of 557,357 entries, 159,669 entries were flagged for house number changes (i.e. 0 or 1) - 28.6%.

7,350 entries were flagged with "1" meaning the house number was editted by the function. E.g. The initial household number might have been "195-7", "34 TO 36" or "112 114". There would be a split between the house numbers and `hn_1` will be 195 and `hn_2` will be 197 as in the first example.

```{r}
# Problematic Enumeration Districts for House Number Cleaning
mn_output %>% filter(flag_hn_cleaned == 1) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(24)
```

Out of all the house hold number entries, 364,730 entries were filled in via the Fill Down function `05_House_Number_Fill_Down` - 65.4%.

***

```{r Bk EDA script, include = FALSE}
# Brooklyn EDA

bk_output

# How many EDs in Brooklyn
bk_output$ED %>% unique() %>% length()

# How many Microfilms in Brooklyn
bk_output$microfilm %>% unique() %>% length()

# How many unique Street Names in Brooklyn
bk_output$street_add %>% unique() %>% length()

# Street Names Brooklyn
bk_output$best_match %>% unique() %>% head(200)

# Total number of entries from sample dataset in Brooklyn
nrow(bk_output)

# Tabulate Result Type in Brooklyn
# 1 of 6 possible match types: (1) Perfect Match, (2) Identical Match, (3) Singular Mode, (4) Multiple Modes, (5) NA mode, (6) No match. Refer to documentation for more details.
table(bk_output$result_type)

# Tabulate Flagged Filled Streets in Brooklyn
table(bk_output$flg_filled_st)

# Tabulate Flagged House Number Cleaned in Brooklyn
table(bk_output$flag_hn_cleaned)

# Tabulate Flagged Filled House Numbers in Brooklyn
table(bk_output$flg_filled_hn)
```

#### Brooklyn

From the sample dataset, an output of 371,833 rows by 33 columns was derived. There are 1,106 Enumeration Districts, 1,527 microfilms and 7,076 unique street names within the dataset.

The results from the Street Matching function are as follow:

- 76.5% Perfect Match
- 7.9% Identical Match
- 9.5% Singular Mode
- 1.5% Multiple Modes
- 1.3% NAs
- 0.8% No Match

In contrast to Manhattan, there is a larger proportion of Perfect Matches.

Out of the 76.5% of Perfect Matches (284k entries) 1,005 entries were matched via the Fill Down function `03_Matched_Street_Fill_Down` - A much smaller percentage than Manhattan. This suggests that the recording of entries might be more accurate in Brooklynn or that the street directory is more well-developed. Nevertheless, errors are still present.

Out of the problematic Street Matches (result type 5 and type 6), there is a trend of some EDs being more problematic than others, i.e. more entries of 5 or 6 within certain Enumeration Districts.

```{r}
# Problematic EDs for Result Type 5
# Treshold set to 50 (arbitrarily decided)
bk_output %>% filter(result_type == 5) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(27)
```

```{r}
# Problematic EDs for Result Type 6
# Treshold set to 50 (arbitrarily decided)
bk_output %>% filter(result_type == 6) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(14)
```

Out of 371,833 entries, 179,303 entries were flagged for house number changes (i.e. 0 or 1) - 48.2%.

2,567 entries were flagged with "1" meaning the house number was editted by the function. Brooklyn also has several abnormal initial house numbers. E.g. "222 1/2", "192 3TH". Otherwise, the issues are similar to Manhattan.

```{r}
# Problematic Enumeration Districts for House Number Cleaning
bk_output %>% filter(flag_hn_cleaned == 1) %>% 
  select(ED) %>% table() %>% sort(decreasing = TRUE) %>% head(11)
```

Out of all the house hold number entries, 182,935 entries were filled in via the Fill Down function `05_House_Number_Fill_Down` - 49.1%.

***

#### Visualizing the Outputs

```{r load GIS libraries, include = FALSE}
p_load(ggmap,
       rgdal,
       rgeos,
       maptools,
       tmap)
```

```{r prepare shapefiles, include = FALSE}
mn_map <- readOGR(dsn = "mn_shapefiles", layer = "Manhattan_1910")
bk_map <- readOGR(dsn = "bk_shapefiles", layer = "Brooklyn_1910")
```

```{r prepare data for MN}
# Aggregate the mean of result type, i.e. street matching success rate
# The lower the better
mn_output_result_type <- aggregate(result_type ~ ED, FUN = mean, data = mn_output)
mn_output_result_type$ED <- as.numeric(mn_output_result_type$ED)

# Change Manhattan Shapefile ED variable to numeric format
mn_map@data$ED <- as.numeric(mn_map@data$ED)

# Left join the 2 datasets
mn_map@data <- left_join(mn_map@data, mn_output_result_type, by = c('ED' = 'ED'))
```

```{r}
# Map for Result Type in Manhattan
tmap_mode("view")
tm_shape(mn_map) + tm_polygons("result_type")
```

```{r prepare data for BK}
# Aggregate the mean of result type, i.e. street matching success rate
# The lower the better
bk_output_result_type <- aggregate(result_type ~ ED, FUN = mean, data = bk_output)
bk_output_result_type$ED <- as.numeric(bk_output_result_type$ED)

# Change Brooklyn Shapefile ED variable to numeric format
bk_map@data$ED <- as.numeric(bk_map@data$ED)

# Left join the 2 datasets
bk_map@data <- left_join(bk_map@data, bk_output_result_type, by = c('ED' = 'ED'))
```

```{r}
# Map for Result Type in Brooklyn
tmap_mode("view")
tm_shape(bk_map) + tm_polygons("result_type")
```

```{r}
# Highlight 2 examples in Brooklyn with poor street name matches
bk_output %>% filter(ED == 1034 | ED == 1039)
```

# Problems to Look into

- Missing EDs in the entire Dataset
    - There are missing EDs in the entire dataset
    - The EDs could be names wrongly
        - In the Shapefile, the max ED is 1480 but the max ED in the output runs till 1941

```{r}
# Missing EDs in Manhattan

# Manhattan ED list
mn_ed_list <- mn_map@data$ED %>% sort()

mn_output_ed <- mn_output$ED %>% unique() %>% as.numeric() %>% sort()
```

```{r}
# Which EDs are in the Shapefile but missing from the Manhattan dataset
mn_ed_list[!(mn_output_ed %in% mn_ed_list)]
length(mn_ed_list[!(mn_output_ed %in% mn_ed_list)])
```

```{r}
# Which EDs are in the dataset but missing from the Manhattan shapefile
mn_output_ed[!(mn_ed_list %in% mn_output_ed)]
length(mn_output_ed[!(mn_ed_list %in% mn_output_ed)])
```

```{r}
# Missing EDs in Brooklyn

# Brooklyn ED list
bk_ed_list <- bk_map@data$ED %>% sort()

# Brooklyn Output EDs
bk_output_ed <- bk_output$ED %>% unique() %>% as.numeric() %>% sort()
```

```{r}
# Which EDs are in the Shapefile but missing from the Brooklyn dataset
bk_ed_list[!(bk_output_ed %in% bk_ed_list)]
length(bk_ed_list[!(bk_output_ed %in% bk_ed_list)])
```

```{r}
# Which EDs are in the dataset but missing from the Manhattan shapefile
bk_output_ed[!(bk_ed_list %in% bk_output_ed)]
length(bk_output_ed[!(bk_ed_list %in% bk_output_ed)])
```

- Why are there missing House Numbers despite the filldown
