---
title: "HNYC Cleaning"
author: "Kyi Yeung Goh"
date: "06/06/2019"
output: html_document
---

```{r setup, include=FALSE}
library(zoo)
library(dplyr)
library(tidyverse)
library(readxl)
library(stringr)
library(fuzzyjoin)
library(reshape2)
df2<-read_csv("us1910m_usa_sample100k.csv")
oldstreets1<-read_xlsx("OldStreets_StreetNames_CleanNotChecked.xlsx")
ed_descriptionMN<-read_xlsx("ED_descriptionsMN1910.xlsx")
manhattan_dan<-read_csv("mn_segments_export.csv")
sm_manhattan<-read_xlsx("SteveMorse_ManhattanStreets.xlsx") #upload
combined_MN_dict<-read_csv("combined_edict_mn.csv")
combined_MN_dict<-combined_MN_dict%>%
  select(-X1)

EDict<-manhattan_dan%>%
  select("streets"=MN_STREET1910_FULL_STREE,"ED1"=MN_STREET1910_MINUS_ED_ED,"ED2"=MN_STREET1910_PLUS_ED_ED)

EDict1<-manhattan_dan%>%
  select("streets"=MN_STREET1910_Y1910NAMESTREET,"ED1"=MN_STREET1910_MINUS_ED_ED,"ED2"=MN_STREET1910_PLUS_ED_ED)

EDict2<-manhattan_dan%>%
  select("streets"=MN_STREET1910_Y1910NAMEALT,"ED1"=MN_STREET1910_MINUS_ED_ED,"ED2"=MN_STREET1910_PLUS_ED_ED)

EDict_MN<-rbind(EDict,EDict1,EDict2)

EDict_MN$cleaned<-EDict_MN$streets
EDict_MN$cleaned<-gsub("\\<\\\\>","",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<SRT\\>|\\<SR\\>\\<SRT\\>|\\<STR\\>|\\<SST\\>|\\<SEET\\>|\\<TREET\\>|\\<SREET\\>|\\<SRT\\>|\\<REET\\>|\\<ST\\>","STREET",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<N\\>","NORTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<S\\>","SOUTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<E\\>","EAST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<W\\>","WEST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<DR\\>|\\<DV\\>|\\<DE\\>|\\<DRV\\>|\\<DRI\\>|\\<DRIV\\>|\\<DRIE\\>","DRIVE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<CIR\\>|\\<CRL\\>|\\<CIRC\\>|\\<CR\\>|\\<CL\\>|\\<CIRCL\\>|\\<CICLE\\>","CIRCLE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<AVE\\>|\\<AV\\>|\\<AVN\\>|\\<AVEN\\>|\\<AVENU\\>","AVENUE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<CT\\>|\\<CRT\\>|\\<CTR\\>|\\<COUR\\>|\\<COT\\>|\\<CORT\\>","COURT",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<BLVD\\>|\\<BVLD\\>|\\<BV\\>|\\<BLD\\>|\\<BD\\>|\\<BL\\>|\\<BLV\\>","BOULEVARD",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<ALY\\>|\\<AL\\>|\\<ALLY\\>|\\<ALEY\\>|\\<ALLE\\>|\\<AY\\>","ALLEY",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<PL\\>|\\<PLC\\>|\\<PLE\\>|\\<PC\\>|\\<PLAC\\>|\\<PLCE\\>|\\<PCE\\>","PLACE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<PKWY\\>|\\<PARKW\\>|\\<PWY\\>|\\<PKW\\>|\\<PRKWY\\>|\\<PKWY\\>|\\<PKW\\>","PARKWAY",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<APPR\\>|\\<APR\\>|\\<APPROA\\>|\\<APRCH\\>|\\<APPRCH\\>","APPROACH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<TER\\>|\\<TERR\\>|\\<TRC\\>|\\<TRCE\\>|\\<TR\\>","TERRACE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<PLZ\\>|\\<PLAZ\\>|\\<PZ\\>|\\<PLZA\\>","PLAZA",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<LN\\>|\\<LNE\\>|\\<LAN\\>","LANE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<BRG\\>|\\<BRGD\\>|\\<BGE\\>","BRIDGE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<HL\\>|\\<HLL\\>|\\<HIL\\>","HILL",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<1ST\\>","FIRST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b1\\b","FIRST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<2ND\\>","SECOND",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b2\\b","SECOND",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<3RD\\>","THIRD",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b3\\b","THIRD",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<4TH\\>","FOURTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b4\\b","FOURTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<5TH\\>","FIFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b5\\b","FIFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<6TH\\>","SIXTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b6\\b","SIXTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<7TH\\>","SEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b7\\b","SEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<8TH\\>","EIGHTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b8\\b","EIGHTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<9TH\\>","NINTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b9\\b","NINTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<10TH\\>","TENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b10\\b","TENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<11TH\\>","ELEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b11\\b","ELEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<12TH\\>","TWELFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b12\\b","TWELFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<13TH\\>","THIRTEENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b13\\b","THIRTEENTH",EDict_MN$cleaned)

EDict_MN$cleaned<-gsub("^\\bST\\b","SAINT", EDict_MN$cleaned)
##EDict_MN$cleaned<-str_remove(EDict_MN$cleaned,"STREET")

EDict_MN$ED1<-as.numeric(as.character(EDict_MN$ED1))
EDict_MN$ED2<-as.numeric(as.character(EDict_MN$ED2))
```

### Addressing issues with addresses 

### Working with the 100k sample

### Fill-down first
```{r}
df2<-read_csv("us1910m_usa_sample100k.csv")
for (i in 1:ncol(df2)){
if (is.na(df2[1,i]) == FALSE ){
  df2[,i]=na.locf(df2[,i])}
  else {
  df2[,i]=df2[,i]
    }
}
```

## Household profiles
```{r}
not_all_na <- function(x) any(!is.na(x))
hhprofiles<-df2%>%
  dplyr::filter(is.na(df2$`Dwelling sequence number`)==TRUE)%>% 
  select_if(not_all_na)
```


### Removing the household and then joining it with persons
```{r}

personprofile<-df2%>%
  filter(is.na(df2$`Consistent historical data person identifier`)==FALSE)

```


```{r}
### Address problems
df2$`Street address 2`<-na.locf(df2$`Street address 2`)

### Inconsistent streetnames
distinct_streets2<-personprofile%>%
  select(`Enumeration district 2`,`House number`,`Street address 2`,`Dwelling serial number`,`Dwelling serial number 2`,`Line Number`,`Line number 2`,`Microfilm page number 3`)%>%
  distinct(`Street address 2`,.keep_all=TRUE)

distinct_streets2$`Enumeration district 2`<-as.numeric(as.character(distinct_streets2$`Enumeration district 2`))

colnames(distinct_streets2)[3]<-"streets"
```

### Function to clean existing streetnames and create a new 'cleaned' column, match using that and then drop it after

```{r}
## Preprocessing strings

rem_dup_word <- function(x){
x <- tolower(x)
paste(unique(trimws(unlist(strsplit(x,split=" ",fixed=F,perl=T)))),collapse = 
" ")
}

str_clean<-function(x){
  x$cleaned<-gsub("\\<SRT\\>$|\\<SR\\>$\\<SRT\\>$|\\<STR\\>$|\\<SST\\>$|\\<SEET\\>$|\\<TREET\\>$|\\<SHEER\\>$|\\<SHEE\\>$|\\<STREE\\>$|\\<SREET\\>$|\\<REET\\>$|\\<STEE\\>$|\\<ST\\>$","STREET",x$cleaned)
  x$cleaned<-gsub("\\<N\\>","NORTH",x$cleaned)
  x$cleaned<-gsub("\\<S\\>","SOUTH",x$cleaned)
  x$cleaned<-gsub("\\<E\\>","EAST",x$cleaned)
  x$cleaned<-gsub("\\<W\\>","WEST",x$cleaned)
  x$cleaned<-gsub("\\<DR\\>|\\<DV\\>|\\<DE\\>$|\\<DRV\\>|\\<DRI\\>|\\<DRIV\\>|\\<DRIE\\>","DRIVE",x$cleaned) 
  x$cleaned<-gsub("\\<CIR\\>|\\<CRL\\>|\\<CIRC\\>|\\<CR\\>|\\<CL\\>|\\<CIRCL\\>|\\<CICLE\\>","CIRCLE",x$cleaned)
  x$cleaned<-gsub("\\<AVE\\>|\\<AV\\>|\\<AVN\\>|\\<AVEN\\>|\\<AVENU\\>","AVENUE",x$cleaned)
  x$cleaned<-gsub("\\<CT\\>|\\<CRT\\>|\\<CTR\\>|\\<COUR\\>|\\<COT\\>|\\<CORT\\>","COURT",x$cleaned)
  x$cleaned<-gsub("\\<BLVD\\>|\\<BVLD\\>|\\<BV\\>|\\<BLD\\>|\\<BD\\>|\\<BL\\>|\\<BLV\\>","BOULEVARD",x$cleaned)
  x$cleaned<-gsub("\\<RD\\>|\\<RAD\\>|\\<ROD\\>","ROAD",x$cleaned)
  x$cleaned<-gsub("\\<ALY\\>|\\<AL\\>|\\<ALLY\\>|\\<ALEY\\>|\\<ALLE\\>|\\<AY\\>","ALLEY",x$cleaned)
  x$cleaned<-gsub("\\<PL\\>|\\<PLC\\>|\\<PLE\\>|\\<PC\\>|\\<PLAC\\>|\\<PLCE\\>|\\<PCE\\>","PLACE",x$cleaned)
  x$cleaned<-gsub("\\<PK\\>|\\<PRK\\>|\\<PRAK\\>|\\<PAK\\>","PARK",x$cleaned)
  x$cleaned<-gsub("\\<PKWY\\>|\\<PARKW\\>|\\<PWY\\>|\\<PKW\\>|\\<PRKWY\\>|\\<PKWY\\>|\\<PKW\\>","PARKWAY",x$cleaned)
  x$cleaned<-gsub("\\<APPR\\>|\\<APR\\>|\\<APPROA\\>|\\<APRCH\\>|\\<APPRCH\\>","APPROACH",x$cleaned)
  x$cleaned<-gsub("\\<TER\\>|\\<TERR\\>|\\<TRC\\>|\\<TRCE\\>|\\<TR\\>","TERRACE",x$cleaned)
  x$cleaned<-gsub("\\<PLZ\\>|\\<PLAZ\\>|\\<PZ\\>|\\<PLZA\\>","PLAZA",x$cleaned)
  x$cleaned<-gsub("\\<LN\\>|\\<LNE\\>|\\<LAN\\>","LANE",x$cleaned)
  x$cleaned<-gsub("\\<BRG\\>|\\<BRGD\\>|\\<BGE\\>","BRIDGE",x$cleaned)
  x$cleaned<-gsub("\\<HL\\>|\\<HLL\\>|\\<HIL\\>","HILL",x$cleaned)
  x$cleaned<-gsub("\\<HTS\\>|\\<HT\\>|\\<HEIGHT\\>|\\<HEGHTS\\>|\\<HHT\\>|\\<HEIGT\\>","HEIGHTS",x$cleaned) 
  x$cleaned<-gsub(".*\\((.*)\\).*", "\\1", x$cleaned)
  x$cleaned<-str_remove(x$cleaned,"STREET")
  x$cleaned<-gsub("\\d+\\ - *\\d*|\\d+\\ TO *\\d*|\\d+\\-\\d*","",x$cleaned) #remove addresses
  

  ## dealing with numbered streets
x$cleaned<-gsub("(\\d)(ST|ND|RD|TH)\\b", "\\1", x$cleaned)
x$cleaned<-str_remove(x$cleaned, "(?<=[0-9])(ST|ND|RD|TH)")
x$cleaned<-gsub("\\<ONE HUNDRED\\>|\\<ONEHUNDRED\\>|\\<HUNDRED\\>|\\<HUDRED\\>|\\<HUNDED\\>","1",x$cleaned) 
x$cleaned<-gsub("\\<TWO HUNDRED\\>|\\<TWOHUNDRED\\>","2",x$cleaned)
x$cleaned<-gsub("-"," ",x$cleaned)
x$cleaned<-gsub("\\<AND\\>"," ",x$cleaned)
x$cleaned<-gsub("&"," ",x$cleaned)
x$cleaned<-gsub("\\<1ST\\>|\\b1\\b","FIRST",x$cleaned)
x$cleaned<-gsub("\\<2ND\\>|\\b2\\b","SECOND",x$cleaned)
x$cleaned<-gsub("\\<3RD\\>|\\b3\\b","THIRD",x$cleaned)
x$cleaned<-gsub("\\<4TH\\>|\\b4\\b","FOURTH",x$cleaned)
x$cleaned<-gsub("\\<5TH\\>|\\b5\\b","FIFTH",x$cleaned)
x$cleaned<-gsub("\\<6TH\\>|\\b6\\b","SIXTH",x$cleaned)
x$cleaned<-gsub("\\<7TH\\>|\\b7\\b","SEVENTH",x$cleaned)
x$cleaned<-gsub("\\<8TH\\>|\\b8\\b","EIGHTH",x$cleaned)
x$cleaned<-gsub("\\<9TH\\>|\\b9\\b","NINTH",x$cleaned)
x$cleaned<-gsub("\\<10TH\\>|\\b10\\b","TENTH",x$cleaned)
x$cleaned<-gsub("\\<11TH\\>|\\b11\\b","ELEVENTH",x$cleaned)
x$cleaned<-gsub("\\<12TH\\>|\\b12\\b","TWELFTH",x$cleaned)
x$cleaned<-gsub("\\<13TH\\>|\\b13\\b","THIRTEENTH",x$cleaned)
x$cleaned<-gsub("\\<14TH\\>|\\b14\\b","FORTEENTH",x$cleaned)
x$cleaned<-gsub("\\<15TH\\>|\\b15\\b","FIFTEENTH",x$cleaned)
x$cleaned<-gsub("\\<16TH\\>|\\b16\\b","SIXTEENTH",x$cleaned)
x$cleaned<-gsub("\\<17TH\\>|\\b17\\b","SEVENTEENTH",x$cleaned)
x$cleaned<-gsub("\\<18TH\\>|\\b18\\b","EIGHTEENTH",x$cleaned)
x$cleaned<-gsub("\\<19TH\\>|\\b19\\b","NINETEENTH",x$cleaned)


x$cleaned<-gsub("\\<TWENTY\\>|\\<TWENTI\\>|\\<TENTI\\>","2",x$cleaned)
x$cleaned<-gsub("\\<THIRTY\\>|\\<THIRTHY\\>|\\<THIRTEY\\>|\\<TIRTY\\>|\\<TRITHY\\>","3",x$cleaned)
x$cleaned<-gsub("\\<FORTY\\>|\\<FOURTY\\>|\\<FOURTHY\\>|\\<FRTY\\>","4",x$cleaned)
x$cleaned<-gsub("\\<FIFTY\\>|\\<FIFTEY\\>|\\<FIFT\\>|\\<FITY\\>|\\<FIFTHY\\>","5",x$cleaned)
x$cleaned<-gsub("\\<SIXTY\\>|\\<SXTY\\>|\\<SIXY\\>|\\<SXTY\\>|\\<SIXTHY\\>|\\<SIXTEY\\>","6",x$cleaned)
x$cleaned<-gsub("\\<SEVENT\\>|\\<SEVENTY\\>|\\<SEVENTEY\\>|\\<SVENTY\\>|\\<SEVENTI\\>","7",x$cleaned)
x$cleaned<-gsub("\\<EIGHTY\\>|\\<EIGHTEY\\>|\\<EIGHTE\\>","8",x$cleaned)
x$cleaned<-gsub("\\<UNITY\\>|\\<NINTH\\>|\\<NINETY\\>|\\<NINETEY\\>|\\<NINETIETH\\>|\\<NINTY\\>","9",x$cleaned)
x$cleaned<-gsub("\\<FRIST\\>|\\<FIST\\>|\\<FRST\\>|\\<FIRST\\>|\\<ONE\\>","1",x$cleaned)
x$cleaned<-gsub("\\<SECOND\\>|\\<SECORD\\>|\\<SCOND\\>|\\<SECOND\\>|\\<TWO\\>","2",x$cleaned)
x$cleaned<-gsub("\\<THRID\\>|\\<THIRD\\>|\\<TIRD\\>|\\<TRIHD\\>|\\<THREE\\>","3",x$cleaned)
x$cleaned<-gsub("\\<FORTH\\>|\\<FOURTH\\>|\\<FROTH\\>|\\<FROUTH\\>|\\<FOUR\\>","4",x$cleaned)
x$cleaned<-gsub("\\<FIFETH\\>|\\<FIFTH\\>|\\<FIFFTH\\>|\\<FIFTHE\\>|\\<FIVE\\>","5",x$cleaned)
x$cleaned<-gsub("\\<SIXTH\\>|\\<SXTH\\>|\\<SITH\\>|\\<SIHXT\\>|\\<SIX\\>","6",x$cleaned)
x$cleaned<-gsub("\\<SEVENTH\\>|\\<SVEN\\>|\\<SVENTH\\>|\\<SEVENH\\>|\\<SEVENT\\>|\\<SEVEN\\>","7",x$cleaned)
x$cleaned<-gsub("\\<EIGHTH\\>|\\<EIGHTEH\\>|\\<EITH\\>|\\<EIGHT\\>","8",x$cleaned)
x$cleaned<-gsub("\\<NINETH\\>|\\<NINTH\\>|\\<NINT\\>|\\<NINETH\\>|\\<NINE\\>|\\<NIN\\>","9",x$cleaned)
x$cleaned<-gsub("\\<TWENTIETH\\>|\\<TWENTIEFTH\\>","20",x$cleaned) #NEW
x$cleaned<-gsub("\\<THIRTIETH\\>|\\<THIRTIEFTH\\>","30",x$cleaned)
x$cleaned<-gsub("\\<FORTIETH\\>|\\<FOURTIETH\\>","40",x$cleaned)
x$cleaned<-gsub("\\<FIFTIETH\\>","50",x$cleaned)
x$cleaned<-gsub("\\<SIXTIETH\\>","60",x$cleaned)
x$cleaned<-gsub("\\<SEVENTIETH\\>","70",x$cleaned)
x$cleaned<-gsub("\\<EIGHTIETH\\>","80",x$cleaned)
x$cleaned<-gsub("\\<NINETIETH\\>|\\<NINTIETH\\>","90",x$cleaned)
x$cleaned<-gsub("(?<=\\d) (?=\\d)","",x$cleaned,perl = T) #new close gaps between all numbers
## place names
  ##x$cleaned<-gsub("\\bSTR\\b","", x$cleaned)
  x$cleaned<-gsub("^\\bST\\b","SAINT", x$cleaned) 
  x$cleaned<-gsub("\\bHOUSE\\b","", x$cleaned)
  x$cleaned<-gsub("\\bHOSTEL\\b","", x$cleaned)
  x$cleaned<-gsub("\\bHOTEL\\b","", x$cleaned)
  x$cleaned<-gsub("\\bLODGE\\b","", x$cleaned)
  x$cleaned<-gsub("\\bLODGING\\b","", x$cleaned)
  x$cleaned<-trimws(x$cleaned, "both")
  x$cleaned<-gsub("\\<N\\>","NORTH",x$cleaned)
  ##x$cleaned<-gsub("\\<ST\\>","",x$cleaned)
  ##x$cleaned<-gsub("\\<STREET\\>","",x$cleaned)
} 

for (i in 1:nrow(distinct_streets2)){
distinct_streets2$cleaned[i]<-toupper(rem_dup_word(distinct_streets2$streets[i]))
}

# create numbered list
distinct_streets2$cleaned<-str_clean(distinct_streets2)
colnames(distinct_streets2)[1]<-"ED"
```

## Reshape original geodata ED dictionary (wide to long)
```{r}
EDict_MN_1<-EDict_MN%>%
  select(-ED2)

colnames(EDict_MN_1)[2]="ED"

EDict_MN_2<-EDict_MN%>%
  select(-ED1)
colnames(EDict_MN_2)[2]="ED"

## long-form of ED1 and ED2 columns

EDict_MN_all<-rbind(EDict_MN_1,EDict_MN_2)

```

## combine with Steve Morse (new column spelling)
```{r}
morse_mn1910<-read_csv("MN_MORSE_EDstreet_dict_1910.csv")
morse_mn_list<-gather(data = morse_mn1910, 
             key = item, 
             value = value, 
             V1:V35)%>%
   select(-item)

morse_mn_list<-na.omit(morse_mn_list)
for (i in 1:nrow(morse_mn_list)){
morse_mn_list$cleaned[i]<-toupper(rem_dup_word(morse_mn_list$value[i]))
}

# create numbered list
morse_mn_list$cleaned<-str_clean(morse_mn_list)
colnames(morse_mn_list)[2]="streets"


```

### Creating column for EDict which gives us name type
```{r}
## combine
EDICT_ALL_MN<-rbind(morse_mn_list,EDict_MN_all)
EDICT_ALL_MN$streets<-toupper(EDICT_ALL_MN$streets)
EDICT_ALL_MN<-EDICT_ALL_MN%>%
  dplyr::filter(cleaned!="NULL")

for (i in 1:nrow(EDICT_ALL_MN)){
if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bSTREET\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"STREET"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bAVENUE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"AVENUE" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bDRIVE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"DRIVE" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bCIRCLE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"CIRCLE" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bCOURT\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"COURT" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bBOULEVARD\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"BOULEVARD" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bALLEY\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"ALLEY"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPLACE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"PLACE" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPARKWAY\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"PARKWAY"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPARK\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"PARK"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bAPPROACH\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"APPROACH"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bTERRACE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"TERRACE" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPLAZA\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"PLAZA" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bLANE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"LANE"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bBRIDGE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"BRIDGE" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bHILL\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"HILL"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bHEIGHTS\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"HEIGHTS" 
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bHOSPITAL\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"HOSPITAL"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bASYLUM\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"ASYLUM"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bISLAND\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"ISLAND"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bJAIL\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"JAIL"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bRIVER\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"RIVER"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bSQUARE\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"SQUARE"
}else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bROAD\\b")==TRUE){
  EDICT_ALL_MN$type[i]<-"ROAD"
} else {
  EDICT_ALL_MN$type[i]<-""
}
}

EDICT_ALL_MN$cleaned<-gsub("\\<STREET\\>","",EDICT_ALL_MN$cleaned) #remove word 'street'
EDICT_ALL_MN$cleaned<-trimws(EDICT_ALL_MN$cleaned,"right")
EDICT_ALL_MN<-EDICT_ALL_MN%>%
  distinct(ED,cleaned,type,.keep_all = TRUE)
EDICT_ALL_MN[EDICT_ALL_MN==""] <- NA

## fill-in matching strings in 'type' column 
EDICT_ALL_MN<-EDICT_ALL_MN %>% 
  dplyr::group_by(cleaned) %>% 
  tidyr::fill(type)%>%
  tidyr::fill(type, .direction = "up")%>%
  distinct(ED,cleaned,type,.keep_all = TRUE)

### Bring anything with direction at end of string to the front (n/s/e/w) + then label them as streets

EDICT_ALL_MN$cleaned<-sub("(.*?)\\s+(NORTH|EAST|WEST|SOUTH)$", "\\2 \\1", EDICT_ALL_MN$cleaned) #make sure that the distinct misspellings are also processed in the same manner 

for (i in 1:nrow(EDICT_ALL_MN)){
if (str_detect(EDICT_ALL_MN$cleaned[i], "^NORTH|^SOUTH|^EAST|^WEST")==TRUE){
  EDICT_ALL_MN$type[i]="STREET"
}
}

for (i in 1:nrow(EDICT_ALL_MN)){
if (str_detect(EDICT_ALL_MN$cleaned[i], "^NORTH|^SOUTH|^EAST|^WEST")==TRUE){
  EDICT_ALL_MN$type[i]="STREET"
}
}

for (i in 1:nrow(EDICT_ALL_MN)){
if (str_detect(EDICT_ALL_MN$cleaned[i], "^(\\d)(ST|ND|RD|TH)$")==TRUE){
  EDICT_ALL_MN$type[i]="STREET"
}
}


write.csv(EDICT_ALL_MN,"full_MN_dictionary.csv")

```

```{r, warning=FALSE}
### first find ED of concern for that observation then search if not existent then search ALL and note this occurence
### search range to extract dictionary

datalist = list()

for (i in 1:nrow(distinct_streets2)){
  
ED_SAMPLE<-EDICT_ALL_MN%>%
  dplyr::filter(ED==distinct_streets2$ED[i])%>%
  dplyr::filter(streets!="NULL")
  
  if (nrow(ED_SAMPLE)!=0){
check<-stringdist_join(distinct_streets2[i,], ED_SAMPLE, 
                           by = "cleaned", #must be named streets
                           mode = "left",
                           ignore_case = FALSE, 
                           method = "jw",  #jaro-winkler distance works best for this set, can be tuned.
                           max_dist = 1, 
                           distance_col = "dscore")%>%
  dplyr::group_by(cleaned.x)%>%
  dplyr::filter(dscore==min(dscore)) %>%
  head(1)

check$ED_correct<-"YES"

datalist[[i]]<-check
  }else{
check<-stringdist_join(distinct_streets2[i,], EDICT_ALL_MN, 
                           by = "cleaned", #must be named streets
                           mode = "left",
                           ignore_case = FALSE, 
                           method = "jw",  #jaro-winkler distance works best for this set, can be tuned.
                           max_dist = 1, 
                           distance_col = "dscore")%>%
  dplyr::group_by(cleaned.x)%>%
  dplyr::filter(dscore==min(dscore)) %>%
  head(1)

check$ED_correct="NO"

datalist[[i]]<-check

  }
} 

address_check<-do.call("rbind", datalist)

##write.csv(address_check,"address_check.csv")

address_check_flags<-address_check%>%
  filter(dscore>0.27)

##write.csv(address_check_flags,"address_check_flags.csv")

address_match<-data.frame(address_check)%>%
  select(streets, ED_streetname=streets.y,corrected_str=cleaned.y,dscore,type)
### if string detect street then duplicate over into clean, positional matchng where it adds word only if some variation of street appears at the back 

for (i in 1:nrow(address_match)){
if(str_detect(address_match$streets.x[i],("\\bSRT$|\\bSR$|\\bSTR$|\\bSST$|\\bSEET$|\\bTREET$|\\bSHEER$|\\bSHEE$|\\bSTREE$|\\bSREET$|\\bREET$|\\bSTEE$|\\bST$|\\bSTREET$"))==TRUE){
  address_match$cleaned.y[i]<-paste(address_match$cleaned.y[i]," STREET",sep="")
}else if(str_detect(address_match$streets.y[i],("\\bSRT$|\\bSR$|\\bSTR$|\\bSST$|\\bSEET$|\\bTREET$|\\bSHEER$|\\bSHEE$|\\bSTREE$|\\bSREET$|\\bREET$|\\bSTEE$|\\bST$|\\bSTREET$"))==TRUE){
  address_match$cleaned.y[i]<-paste(address_match$cleaned.y[i]," STREET",sep="")
}else{
  address_match$cleaned.y[i]<-address_match$cleaned.y[i]
}
}


## match based on original street address names column
colnames(df2)[22]="streets" ## both need to have the same column names
colnames(address_match)[1]="streets"

## original address columns are listed as "streets"
## address names in the dictionary are listed as "streets.x"
## matched address is listed as "cleaned.y"

df2_address<-left_join(df2,address_match,by="streets")

write_csv(df2_address,"100ksample_MN_matched.csv")

```



