---
title: "HNYC Cleaning"
author: "Bo Jumrustanasan"
date: "11/6/19"
output: html_document
editor_options: 
  chunk_output_type: console
---

# 1. This could be integrated into the documentation Jolene's created
# 2. Code was cleaned and refactored. The only jobs left here are 
## (1) to have systematic functions for street cleaning and integrate them into the work flow.
## (2) to match hh profiles with cleaned streets back to person profiles

```{r setup, include=FALSE}
library(zoo)
library(tidyverse)
library(readxl)
library(stringr)
library(fuzzyjoin)
library(reshape2)
```

```{r, eval=FALSE}
oldstreets1<-read_xlsx("OldStreets_StreetNames_CleanNotChecked.xlsx")
ed_descriptionMN<-read_xlsx("ED_descriptionsMN1910.xlsx")
manhattan_dan<-read_csv("mn_segments_export.csv")
sm_manhattan<-read_xlsx("SteveMorse_ManhattanStreets.xlsx") #upload
combined_MN_dict<-read_csv("combined_edict_mn.csv")
combined_MN_dict<-combined_MN_dict%>%
  select(-X1)

EDict<-manhattan_dan%>%
  select("streets"=MN_STREET1910_FULL_STREE,"ED1"=MN_STREET1910_MINUS_ED_ED,"ED2"=MN_STREET1910_PLUS_ED_ED)

EDict1<-manhattan_dan%>%
  select("streets"=MN_STREET1910_Y1910NAMESTREET,"ED1"=MN_STREET1910_MINUS_ED_ED,"ED2"=MN_STREET1910_PLUS_ED_ED)

EDict2<-manhattan_dan%>%
  select("streets"=MN_STREET1910_Y1910NAMEALT,"ED1"=MN_STREET1910_MINUS_ED_ED,"ED2"=MN_STREET1910_PLUS_ED_ED)

EDict_MN<-rbind(EDict,EDict1,EDict2)

EDict_MN$cleaned<-EDict_MN$streets
EDict_MN$cleaned<-gsub("\\<\\\\>","",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<SRT\\>|\\<SR\\>\\<SRT\\>|\\<STR\\>|\\<SST\\>|\\<SEET\\>|\\<TREET\\>|\\<SREET\\>|\\<SRT\\>|\\<REET\\>|\\<ST\\>","STREET",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<N\\>","NORTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<S\\>","SOUTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<E\\>","EAST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<W\\>","WEST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<DR\\>|\\<DV\\>|\\<DE\\>|\\<DRV\\>|\\<DRI\\>|\\<DRIV\\>|\\<DRIE\\>","DRIVE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<CIR\\>|\\<CRL\\>|\\<CIRC\\>|\\<CR\\>|\\<CL\\>|\\<CIRCL\\>|\\<CICLE\\>","CIRCLE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<AVE\\>|\\<AV\\>|\\<AVN\\>|\\<AVEN\\>|\\<AVENU\\>","AVENUE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<CT\\>|\\<CRT\\>|\\<CTR\\>|\\<COUR\\>|\\<COT\\>|\\<CORT\\>","COURT",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<BLVD\\>|\\<BVLD\\>|\\<BV\\>|\\<BLD\\>|\\<BD\\>|\\<BL\\>|\\<BLV\\>","BOULEVARD",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<ALY\\>|\\<AL\\>|\\<ALLY\\>|\\<ALEY\\>|\\<ALLE\\>|\\<AY\\>","ALLEY",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<PL\\>|\\<PLC\\>|\\<PLE\\>|\\<PC\\>|\\<PLAC\\>|\\<PLCE\\>|\\<PCE\\>","PLACE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<PKWY\\>|\\<PARKW\\>|\\<PWY\\>|\\<PKW\\>|\\<PRKWY\\>|\\<PKWY\\>|\\<PKW\\>","PARKWAY",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<APPR\\>|\\<APR\\>|\\<APPROA\\>|\\<APRCH\\>|\\<APPRCH\\>","APPROACH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<TER\\>|\\<TERR\\>|\\<TRC\\>|\\<TRCE\\>|\\<TR\\>","TERRACE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<PLZ\\>|\\<PLAZ\\>|\\<PZ\\>|\\<PLZA\\>","PLAZA",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<LN\\>|\\<LNE\\>|\\<LAN\\>","LANE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<BRG\\>|\\<BRGD\\>|\\<BGE\\>","BRIDGE",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<HL\\>|\\<HLL\\>|\\<HIL\\>","HILL",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<1ST\\>","FIRST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b1\\b","FIRST",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<2ND\\>","SECOND",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b2\\b","SECOND",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<3RD\\>","THIRD",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b3\\b","THIRD",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<4TH\\>","FOURTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b4\\b","FOURTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<5TH\\>","FIFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b5\\b","FIFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<6TH\\>","SIXTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b6\\b","SIXTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<7TH\\>","SEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b7\\b","SEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<8TH\\>","EIGHTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b8\\b","EIGHTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<9TH\\>","NINTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b9\\b","NINTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<10TH\\>","TENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b10\\b","TENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<11TH\\>","ELEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b11\\b","ELEVENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<12TH\\>","TWELFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b12\\b","TWELFTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\<13TH\\>","THIRTEENTH",EDict_MN$cleaned)
EDict_MN$cleaned<-gsub("\\b13\\b","THIRTEENTH",EDict_MN$cleaned)

EDict_MN$cleaned<-gsub("^\\bST\\b","SAINT", EDict_MN$cleaned)
##EDict_MN$cleaned<-str_remove(EDict_MN$cleaned,"STREET")

EDict_MN$ED1<-as.numeric(as.character(EDict_MN$ED1))
EDict_MN$ED2<-as.numeric(as.character(EDict_MN$ED2))
```

# Matching Streets from Census Data to Streets in ED Dictionary

This file walks through the process of matching streets from census data to corrected streets from the street dictionary.

## 1.Census Data Preperation

### 1.1 Read in sample census data
```{r}
## change the path to your relative file directory
df2 <- read_csv("/Users/panchanok/Desktop/HNYC/data/us1910m_usa_sample100k.csv")
```

### 1.2 Creating dataframe of distinct addresses from sample census data

Since 'person' records under the same household record have the same address, it is more efficieint to match only streets in household records and merge the result back to person records at the end. Notice that household records do not have page numbers. Before the estraction, we infer the page number by filling in with the page number of a person record immediately below the household record. With the same logic, it is even more efficient to keep only 1 instance for each unique address. Only address columns are kept:

* `Enumeration district 2`
* `House number`
* `Street address 2` (This column sis renamed to "streets")
* `Dwelling serial number`
* `Dwelling serial number 2`
* `Line Number`
* `Line number 2`
* `Microfilm page number 3`

```{r}
## some tests needed here

## checked: all h records have `Street address 2`
hh_profiles <- df2 %>% 
  fill(`Microfilm page number 3`, .direction = "up") %>%
  filter(`Record type` == "H") %>%
  dplyr::select(`Enumeration district 2`,`House number`,`Street address 2`,
                `Dwelling serial number`,`Dwelling serial number 2`,`Line Number`,
                `Line number 2`,`Microfilm page number 3`)%>%
  distinct(`Street address 2`,.keep_all=TRUE) %>%
  mutate(`Enumeration district 2` = as.numeric(`Enumeration district 2`)) %>%
  rename(streets = `Street address 2`)
```

### 1.3 Create a function that cleans/standardizes addresses

There are some symbols and spaces that could deteriorate the address matching process. The function created at this step removes those symbols and standardizes the addresses:

[Cleaning decription here]

### These are functions for street cleaning
```{r}

rem_dup_word <- function(x){
  x <- tolower(x)
  paste(unique(trimws(unlist(strsplit(x,split=" ",fixed=F,perl=T)))),collapse = " ")
}

### Function to clean existing streetnames and create a new 'cleaned' column, match using that and then drop it after
str_clean<-function(x){
  x$cleaned<-gsub("\\<SRT\\>$|\\<SR\\>$\\<SRT\\>$|\\<STR\\>$|\\<SST\\>$|\\<SEET\\>$|\\<TREET\\>$|\\<SHEER\\>$|\\<SHEE\\>$|\\<STREE\\>$|\\<SREET\\>$|\\<REET\\>$|\\<STEE\\>$|\\<ST\\>$","STREET",x$cleaned)
  x$cleaned<-gsub("\\<N\\>","NORTH",x$cleaned)
  x$cleaned<-gsub("\\<S\\>","SOUTH",x$cleaned)
  x$cleaned<-gsub("\\<E\\>","EAST",x$cleaned)
  x$cleaned<-gsub("\\<W\\>","WEST",x$cleaned)
  x$cleaned<-gsub("\\<DR\\>|\\<DV\\>|\\<DE\\>$|\\<DRV\\>|\\<DRI\\>|\\<DRIV\\>|\\<DRIE\\>","DRIVE",x$cleaned) 
  x$cleaned<-gsub("\\<CIR\\>|\\<CRL\\>|\\<CIRC\\>|\\<CR\\>|\\<CL\\>|\\<CIRCL\\>|\\<CICLE\\>","CIRCLE",x$cleaned)
  x$cleaned<-gsub("\\<AVE\\>|\\<AV\\>|\\<AVN\\>|\\<AVEN\\>|\\<AVENU\\>","AVENUE",x$cleaned)
  x$cleaned<-gsub("\\<CT\\>|\\<CRT\\>|\\<CTR\\>|\\<COUR\\>|\\<COT\\>|\\<CORT\\>","COURT",x$cleaned)
  x$cleaned<-gsub("\\<BLVD\\>|\\<BVLD\\>|\\<BV\\>|\\<BLD\\>|\\<BD\\>|\\<BL\\>|\\<BLV\\>","BOULEVARD",x$cleaned)
  x$cleaned<-gsub("\\<RD\\>|\\<RAD\\>|\\<ROD\\>","ROAD",x$cleaned)
  x$cleaned<-gsub("\\<ALY\\>|\\<AL\\>|\\<ALLY\\>|\\<ALEY\\>|\\<ALLE\\>|\\<AY\\>","ALLEY",x$cleaned)
  x$cleaned<-gsub("\\<PL\\>|\\<PLC\\>|\\<PLE\\>|\\<PC\\>|\\<PLAC\\>|\\<PLCE\\>|\\<PCE\\>","PLACE",x$cleaned)
  x$cleaned<-gsub("\\<PK\\>|\\<PRK\\>|\\<PRAK\\>|\\<PAK\\>","PARK",x$cleaned)
  x$cleaned<-gsub("\\<PKWY\\>|\\<PARKW\\>|\\<PWY\\>|\\<PKW\\>|\\<PRKWY\\>|\\<PKWY\\>|\\<PKW\\>","PARKWAY",x$cleaned)
  x$cleaned<-gsub("\\<APPR\\>|\\<APR\\>|\\<APPROA\\>|\\<APRCH\\>|\\<APPRCH\\>","APPROACH",x$cleaned)
  x$cleaned<-gsub("\\<TER\\>|\\<TERR\\>|\\<TRC\\>|\\<TRCE\\>|\\<TR\\>","TERRACE",x$cleaned)
  x$cleaned<-gsub("\\<PLZ\\>|\\<PLAZ\\>|\\<PZ\\>|\\<PLZA\\>","PLAZA",x$cleaned)
  x$cleaned<-gsub("\\<LN\\>|\\<LNE\\>|\\<LAN\\>","LANE",x$cleaned)
  x$cleaned<-gsub("\\<BRG\\>|\\<BRGD\\>|\\<BGE\\>","BRIDGE",x$cleaned)
  x$cleaned<-gsub("\\<HL\\>|\\<HLL\\>|\\<HIL\\>","HILL",x$cleaned)
  x$cleaned<-gsub("\\<HTS\\>|\\<HT\\>|\\<HEIGHT\\>|\\<HEGHTS\\>|\\<HHT\\>|\\<HEIGT\\>","HEIGHTS",x$cleaned) 
  x$cleaned<-gsub(".*\\((.*)\\).*", "\\1", x$cleaned)
  x$cleaned<-str_remove(x$cleaned,"STREET")
  x$cleaned<-gsub("\\d+\\ - *\\d*|\\d+\\ TO *\\d*|\\d+\\-\\d*","",x$cleaned) #remove addresses
  

  ## dealing with numbered streets
x$cleaned<-gsub("(\\d)(ST|ND|RD|TH)\\b", "\\1", x$cleaned)
x$cleaned<-str_remove(x$cleaned, "(?<=[0-9])(ST|ND|RD|TH)")
x$cleaned<-gsub("\\<ONE HUNDRED\\>|\\<ONEHUNDRED\\>|\\<HUNDRED\\>|\\<HUDRED\\>|\\<HUNDED\\>","1",x$cleaned) 
x$cleaned<-gsub("\\<TWO HUNDRED\\>|\\<TWOHUNDRED\\>","2",x$cleaned)
x$cleaned<-gsub("-"," ",x$cleaned)
x$cleaned<-gsub("\\<AND\\>"," ",x$cleaned)
x$cleaned<-gsub("&"," ",x$cleaned)
x$cleaned<-gsub("\\<1ST\\>|\\b1\\b","FIRST",x$cleaned)
x$cleaned<-gsub("\\<2ND\\>|\\b2\\b","SECOND",x$cleaned)
x$cleaned<-gsub("\\<3RD\\>|\\b3\\b","THIRD",x$cleaned)
x$cleaned<-gsub("\\<4TH\\>|\\b4\\b","FOURTH",x$cleaned)
x$cleaned<-gsub("\\<5TH\\>|\\b5\\b","FIFTH",x$cleaned)
x$cleaned<-gsub("\\<6TH\\>|\\b6\\b","SIXTH",x$cleaned)
x$cleaned<-gsub("\\<7TH\\>|\\b7\\b","SEVENTH",x$cleaned)
x$cleaned<-gsub("\\<8TH\\>|\\b8\\b","EIGHTH",x$cleaned)
x$cleaned<-gsub("\\<9TH\\>|\\b9\\b","NINTH",x$cleaned)
x$cleaned<-gsub("\\<10TH\\>|\\b10\\b","TENTH",x$cleaned)
x$cleaned<-gsub("\\<11TH\\>|\\b11\\b","ELEVENTH",x$cleaned)
x$cleaned<-gsub("\\<12TH\\>|\\b12\\b","TWELFTH",x$cleaned)
x$cleaned<-gsub("\\<13TH\\>|\\b13\\b","THIRTEENTH",x$cleaned)
x$cleaned<-gsub("\\<14TH\\>|\\b14\\b","FORTEENTH",x$cleaned)
x$cleaned<-gsub("\\<15TH\\>|\\b15\\b","FIFTEENTH",x$cleaned)
x$cleaned<-gsub("\\<16TH\\>|\\b16\\b","SIXTEENTH",x$cleaned)
x$cleaned<-gsub("\\<17TH\\>|\\b17\\b","SEVENTEENTH",x$cleaned)
x$cleaned<-gsub("\\<18TH\\>|\\b18\\b","EIGHTEENTH",x$cleaned)
x$cleaned<-gsub("\\<19TH\\>|\\b19\\b","NINETEENTH",x$cleaned)


x$cleaned<-gsub("\\<TWENTY\\>|\\<TWENTI\\>|\\<TENTI\\>","2",x$cleaned)
x$cleaned<-gsub("\\<THIRTY\\>|\\<THIRTHY\\>|\\<THIRTEY\\>|\\<TIRTY\\>|\\<TRITHY\\>","3",x$cleaned)
x$cleaned<-gsub("\\<FORTY\\>|\\<FOURTY\\>|\\<FOURTHY\\>|\\<FRTY\\>","4",x$cleaned)
x$cleaned<-gsub("\\<FIFTY\\>|\\<FIFTEY\\>|\\<FIFT\\>|\\<FITY\\>|\\<FIFTHY\\>","5",x$cleaned)
x$cleaned<-gsub("\\<SIXTY\\>|\\<SXTY\\>|\\<SIXY\\>|\\<SXTY\\>|\\<SIXTHY\\>|\\<SIXTEY\\>","6",x$cleaned)
x$cleaned<-gsub("\\<SEVENT\\>|\\<SEVENTY\\>|\\<SEVENTEY\\>|\\<SVENTY\\>|\\<SEVENTI\\>","7",x$cleaned)
x$cleaned<-gsub("\\<EIGHTY\\>|\\<EIGHTEY\\>|\\<EIGHTE\\>","8",x$cleaned)
x$cleaned<-gsub("\\<UNITY\\>|\\<NINTH\\>|\\<NINETY\\>|\\<NINETEY\\>|\\<NINETIETH\\>|\\<NINTY\\>","9",x$cleaned)
x$cleaned<-gsub("\\<FRIST\\>|\\<FIST\\>|\\<FRST\\>|\\<FIRST\\>|\\<ONE\\>","1",x$cleaned)
x$cleaned<-gsub("\\<SECOND\\>|\\<SECORD\\>|\\<SCOND\\>|\\<SECOND\\>|\\<TWO\\>","2",x$cleaned)
x$cleaned<-gsub("\\<THRID\\>|\\<THIRD\\>|\\<TIRD\\>|\\<TRIHD\\>|\\<THREE\\>","3",x$cleaned)
x$cleaned<-gsub("\\<FORTH\\>|\\<FOURTH\\>|\\<FROTH\\>|\\<FROUTH\\>|\\<FOUR\\>","4",x$cleaned)
x$cleaned<-gsub("\\<FIFETH\\>|\\<FIFTH\\>|\\<FIFFTH\\>|\\<FIFTHE\\>|\\<FIVE\\>","5",x$cleaned)
x$cleaned<-gsub("\\<SIXTH\\>|\\<SXTH\\>|\\<SITH\\>|\\<SIHXT\\>|\\<SIX\\>","6",x$cleaned)
x$cleaned<-gsub("\\<SEVENTH\\>|\\<SVEN\\>|\\<SVENTH\\>|\\<SEVENH\\>|\\<SEVENT\\>|\\<SEVEN\\>","7",x$cleaned)
x$cleaned<-gsub("\\<EIGHTH\\>|\\<EIGHTEH\\>|\\<EITH\\>|\\<EIGHT\\>","8",x$cleaned)
x$cleaned<-gsub("\\<NINETH\\>|\\<NINTH\\>|\\<NINT\\>|\\<NINETH\\>|\\<NINE\\>|\\<NIN\\>","9",x$cleaned)
x$cleaned<-gsub("\\<TWENTIETH\\>|\\<TWENTIEFTH\\>","20",x$cleaned) #NEW
x$cleaned<-gsub("\\<THIRTIETH\\>|\\<THIRTIEFTH\\>","30",x$cleaned)
x$cleaned<-gsub("\\<FORTIETH\\>|\\<FOURTIETH\\>","40",x$cleaned)
x$cleaned<-gsub("\\<FIFTIETH\\>","50",x$cleaned)
x$cleaned<-gsub("\\<SIXTIETH\\>","60",x$cleaned)
x$cleaned<-gsub("\\<SEVENTIETH\\>","70",x$cleaned)
x$cleaned<-gsub("\\<EIGHTIETH\\>","80",x$cleaned)
x$cleaned<-gsub("\\<NINETIETH\\>|\\<NINTIETH\\>","90",x$cleaned)
x$cleaned<-gsub("(?<=\\d) (?=\\d)","",x$cleaned,perl = T) #new close gaps between all numbers
## place names
  ##x$cleaned<-gsub("\\bSTR\\b","", x$cleaned)
  x$cleaned<-gsub("^\\bST\\b","SAINT", x$cleaned) 
  x$cleaned<-gsub("\\bHOUSE\\b","", x$cleaned)
  x$cleaned<-gsub("\\bHOSTEL\\b","", x$cleaned)
  x$cleaned<-gsub("\\bHOTEL\\b","", x$cleaned)
  x$cleaned<-gsub("\\bLODGE\\b","", x$cleaned)
  x$cleaned<-gsub("\\bLODGING\\b","", x$cleaned)
  x$cleaned<-trimws(x$cleaned, "both")
  x$cleaned<-gsub("\\<N\\>","NORTH",x$cleaned)
  ##x$cleaned<-gsub("\\<ST\\>","",x$cleaned)
  ##x$cleaned<-gsub("\\<STREET\\>","",x$cleaned)
} 
```


### 1.4 Applying the cleaning function on the dataframe.

A new column of cleaned addresses is appended to the dataframe.

```{r}
clean_hh_profiles <- hh_profiles %>%
  rowwise() %>%
  str_clean() # <<<<< add function that cleans addresses here

# for (i in 1:nrow(distinct_streets2)){
# distinct_streets2$cleaned[i]<-toupper(rem_dup_word(distinct_streets2$streets[i]))
# }
# 
# # create numbered list
# distinct_streets2$cleaned<-str_clean(distinct_streets2)
# colnames(distinct_streets2)[1]<-"ED"
```



## 2. 1910 geodata and Steve Morse's ED

There are two sources of street referrences: a pull from 1910 geodata and ED information from Steve Morse's

### 2.1 Read in and Reshape Edict dataframe

```{r}
manhattan_dan <- read_csv("/Users/panchanok/Desktop/HNYC/_working/mn_segments_export.csv")
```

```{r, warning=FALSE}
## the new method produces an equivalent df to the old method's
EDict_MN_all <- manhattan_dan %>%
  rename(ED1 = MN_STREET1910_MINUS_ED_ED, ED2=MN_STREET1910_PLUS_ED_ED) %>%
  dplyr::select(MN_STREET1910_FULL_STREE, MN_STREET1910_Y1910NAMESTREET, MN_STREET1910_Y1910NAMEALT, ED1, ED2) %>%
  gather(key=street_sorce, value=streets,-ED1, -ED2) %>%
  select(streets, ED1, ED2) %>%
  gather(key=ed_type, value = ED,-streets) %>%
  select(streets, ED)

head(EDict_MN)
```

### 2.2 Read in and Reshape Steve Morse ED

```{r}
## the new method produces an equivalent df to the old method's
morse_mn1910<-read_csv("/Users/panchanok/Desktop/HNYC/_working/MN_MORSE_EDstreet_dict_1910.csv")
morse_mn_list <- morse_mn1910 %>% 
  gather(key=item, value = streets, V1:V35) %>% 
  dplyr::select(-item) %>% 
  filter(!is.na(streets))
```

### 2.3 Append `Cleaned` column for Clean Steve Morse ED streets
```{r}

## this could be improved
for (i in 1:nrow(morse_mn_list)){
  morse_mn_list$cleaned[i]<-toupper(rem_dup_word(morse_mn_list$value[i]))
}

# create numbered list
morse_mn_list$cleaned<-str_clean(morse_mn_list)
colnames(morse_mn_list)[2]="streets"
```


### 2.4 Rbind Edict_MN_all and morse_mn_list

!!!! required cleaned data onwards
```{r}
## combine
EDICT_ALL_MN<-morse_mn_list %>% 
  rbind(EDict_MN_all) %>%
  mutate(streets = toupper(streets)) %>%
  filter(cleaned != "NULL") %>%
  dplyr::mutate(type=cleaned)


for (i in 1:nrow(EDICT_ALL_MN)){
  if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bSTREET\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"STREET"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bAVENUE\\b|\\bAvenue\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"AVENUE" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bDRIVE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"DRIVE" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bCIRCLE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"CIRCLE" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bCOURT\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"COURT" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bBOULEVARD\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"BOULEVARD" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bALLEY\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"ALLEY"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPLACE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"PLACE" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPARKWAY\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"PARKWAY"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPARK\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"PARK"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bAPPROACH\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"APPROACH"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bTERRACE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"TERRACE" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPLAZA\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"PLAZA" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bLANE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"LANE"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bBRIDGE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"BRIDGE" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bHILL\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"HILL"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bHEIGHTS\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"HEIGHTS" 
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bHOSPITAL\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"HOSPITAL"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bASYLUM\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"ASYLUM"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bISLAND\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"ISLAND"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bJAIL\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"JAIL"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bRIVER\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"RIVER"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bSQUARE\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"SQUARE"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bROAD\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"ROAD"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bSLIP\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"SLIP"
  }else if (str_detect(EDICT_ALL_MN$cleaned[i],"\\bPIER\\b|\\bPIERS\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"PIER"
  }else if(str_detect(EDICT_ALL_MN$streets[i],"\\bST\\b|\\bSTR\\b")==TRUE){
    EDICT_ALL_MN$type[i]<-"STREET"
  } else {
    EDICT_ALL_MN$type[i]<-""
  }
}

EDICT_ALL_MN$cleaned<-gsub("\\<STREET\\>","",EDICT_ALL_MN$cleaned) #remove word 'street'
EDICT_ALL_MN$cleaned<-trimws(EDICT_ALL_MN$cleaned,"right")
EDICT_ALL_MN<-EDICT_ALL_MN%>%
  distinct(ED,cleaned,type,.keep_all = TRUE)
EDICT_ALL_MN[EDICT_ALL_MN==""] <- NA

## fill-in matching strings in 'type' column 
EDICT_ALL_MN<-EDICT_ALL_MN %>% 
  dplyr::group_by(cleaned) %>% 
  tidyr::fill(type)%>%
  tidyr::fill(type, .direction = "up")%>%
  distinct(ED,cleaned,type,.keep_all = TRUE)

### Bring anything with direction at end of string to the front (n/s/e/w) + then label them as streets

EDICT_ALL_MN$cleaned<-sub("(.*?)\\s+(NORTH|EAST|WEST|SOUTH)$", "\\2 \\1", EDICT_ALL_MN$cleaned) #make sure that the distinct misspellings are also processed in the same manner 

for (i in 1:nrow(EDICT_ALL_MN)){
if (str_detect(EDICT_ALL_MN$cleaned[i], "^NORTH|^SOUTH|^EAST|^WEST")==TRUE){
  EDICT_ALL_MN$type[i]="STREET"
}
}

for (i in 1:nrow(EDICT_ALL_MN)){
if (str_detect(EDICT_ALL_MN$cleaned[i], "^(\\d)(ST|ND|RD|TH)$")==TRUE){
  EDICT_ALL_MN$type[i]="STREET"
}
}


write.csv(EDICT_ALL_MN,"full_MN_dictionary.csv")

```

```{r, warning=FALSE}
### first find ED of concern for that observation then search if not existent then search ALL and note this occurence
### search range to extract dictionary

datalist = list()

for (i in 1:nrow(distinct_streets2)){
  
ED_SAMPLE<-EDICT_ALL_MN%>%
  dplyr::filter(ED==distinct_streets2$ED[i])%>%
  dplyr::filter(streets!="NULL")
  
  if (nrow(ED_SAMPLE)!=0){
check<-stringdist_join(distinct_streets2[i,], ED_SAMPLE, 
                           by = "cleaned", #must be named streets
                           mode = "left",
                           ignore_case = FALSE, 
                           method = "jw",  #jaro-winkler distance works best for this set, can be tuned.
                           max_dist = 1, 
                           distance_col = "dscore")%>%
  dplyr::group_by(cleaned.x)%>%
  dplyr::filter(dscore==min(dscore)) %>%
  head(1)

check$ED_correct<-"YES"

datalist[[i]]<-check
  }else{
check<-stringdist_join(distinct_streets2[i,], EDICT_ALL_MN, 
                           by = "cleaned", #must be named streets
                           mode = "left",
                           ignore_case = FALSE, 
                           method = "jw",  #jaro-winkler distance works best for this set, can be tuned.
                           max_dist = 1, 
                           distance_col = "dscore")%>%
  dplyr::group_by(cleaned.x)%>%
  dplyr::filter(dscore==min(dscore)) %>%
  head(1)

check$ED_correct="NO"

datalist[[i]]<-check

  }
} 

address_check<-do.call("rbind", datalist)

##write.csv(address_check,"address_check.csv")

address_check_flags<-address_check%>%
  dplyr::filter(dscore>0.27)

##write.csv(address_check_flags,"address_check_flags.csv")

address_match<-data.frame(address_check)%>%
  dplyr::select(streets.x, ED_streetname=streets.y,corrected_str=cleaned.y,dscore,type)

### if string detect street then duplicate over into clean, positional matchng where it adds word only if some variation of street appears at the back 
address_match$type[is.na(address_match$type)] <- "NA"
for (i in 1:nrow(address_match)){
if(str_detect(address_match$type[i],"STREET")==TRUE){
  address_match$corrected_str[i]<-paste(address_match$corrected_str[i]," STREET",sep="")
}else if(str_detect(address_match$type[i],"\\bNA\\b")==TRUE){
  address_match$corrected_str[i]<-paste(address_match$corrected_str[i]," STREET",sep="")
}else{
  address_match$corrected_str[i]<-address_match$corrected_str[i]
}
}


## match based on original street address names column
colnames(df2)[22]="streets" ## both need to have the same column names
colnames(address_match)[1]="streets"

## original address columns are listed as "streets"
## address names in the dictionary are listed as "streets.x"
## matched address is listed as "cleaned.y"

df2_address<-left_join(df2,address_match,by="streets")

write_csv(df2_address,"100ksample_MN_matched.csv")

```



