---
title: "Full Run of the HNYC Street Cleaning Scripts"
author: "Gerald"
subtitle: Manhattan
output:
  html_document:
    df_print: paged
---
### Set Up
***
#### Import all packages
```{r setup}
library(hunspell)
library(tidyverse)
library(fuzzyjoin)
```

#### Clean House Number Dictionary Function (Step 4A)
```{r 04_1_clean_hn_dict}
clean_hn_dict <- function(file) {
  
  fix_order <- function(str) {
    if (str == "NA?NA") {return (NA_character_)}
    
    str_split(str, "\\?") %>% 
      unlist() %>% 
      as.integer() %>%
      sort() %>% 
      str_c(collapse = "?")
  }
  
  hn_dict <- hn_dict %>%
    mutate(ED = str_pad(ED, 4, "left", pad = "0")) %>%
    select(ED, Name, Left_Low, Left_High, Right_Low, Right_High) %>%
    unite(col = "Left", Left_Low, Left_High, sep = "?") %>% 
    unite(col = "Right", Right_Low, Right_High, sep = "?") %>%
    rowwise() %>%
    mutate(Left = fix_order(Left), Right = fix_order(Right)) %>%
    ungroup() %>%
    gather(key = "column", value = "value", -c(ED, Name)) %>% 
    separate(col = value, into = c("Low", "High"), sep = "\\?") %>%
    mutate(Low = as.integer(Low), High = as.integer(High))
}
```

#### Import data
```{r import data, warning = FALSE}
edict <- read_csv("~/Desktop/HNYC_census_processing/hnyc_census-master/Data/full_mn_dict.csv")
sample <- read_csv("~/Desktop/HNYC_census_processing/hnyc_census-master/Data/census_1910_h_mn.csv")
hn_dict <- read_csv("~/Desktop/HNYC_census_processing/hnyc_census-master/Data/combine_mn.csv")
hn_dict <- hn_dict %>% clean_hn_dict()
```

#### Clean Name of Columns
```{r}
## clean column names: adapt code if sample has different columns. At minimum, ensure `record`, `ED` and `street_add` columns exist after cleaning.
names(sample) <- c("record", "township", "county", "ED", "person_id",
                   "dwelling_seq", "dwelling_ser", "dwelling_ser2", "hh_seq", "hh_seq_8",
                   "hh_ser2", "hh_ser_bef_split", "indiv_seq", "split", 
                   "line_no", "line_no2", "microfilm",
                   "n_fam", "n_person_bef_split", "house_num", "street_add")
```

<br>

### Load all functions
***
#### Street Clean Function `01_Street_Clean_Function_MNBK`
This function preprocesses the `edict`. It will be used in the `02_Street_Matching_MNBK` function.
```{r street clean}
# Updated on 27th February
clean<-function(x){
  x<-gsub("\\<STREET\\>|\\<SRT\\>$|\\<SR\\>$\\<SRT\\>$|\\<STR\\>$|\\<SST\\>$|\\<SEET\\>$|\\<TREET\\>$|\\<SHEER\\>$|\\<SHEE\\>$|\\<STREE\\>$|\\<SREET\\>$|\\<REET\\>$|\\<STEE\\>$|\\<ST\\>$","ST",x)
  x<-gsub("\\<N\\>","N",x)
  x<-gsub("\\<S\\>","S",x)
  x<-gsub("\\<E\\>","E",x)
  x<-gsub("\\<W\\>","W",x)
  x<-gsub("\\<DRIVE\\>|\\<DR\\>|\\<DV\\>|\\<DE\\>$|\\<DRV\\>|\\<DRI\\>|\\<DRIV\\>|\\<DRIE\\>","DR",x) 
  x<-gsub("\\<CIRCLE\\>|\\<CIR\\>|\\<CRL\\>|\\<CIRC\\>|\\<CR\\>|\\<CL\\>|\\<CIRCL\\>|\\<CICLE\\>","CIR",x)
  x<-gsub("\\<AVENUE\\>|\\<AVE\\>|\\<AV\\>|\\<AVN\\>|\\<AVEN\\>|\\<AVENU\\>","AVE",x)
  x<-gsub("\\<COURT\\>|\\<CT\\>|\\<CRT\\>|\\<CTR\\>|\\<COUR\\>|\\<COT\\>|\\<CORT\\>","CT",x)
  x<-gsub("\\<BOULEVARD\\>|\\<BLVD\\>|\\<BVLD\\>|\\<BV\\>|\\<BLD\\>|\\<BD\\>|\\<BL\\>|\\<BLV\\>","BLVD",x)
  x<-gsub("\\<ROAD\\>|\\<RD\\>|\\<RAD\\>|\\<ROD\\>","RD",x)
  x<-gsub("\\<ALLEY\\>|\\<ALY\\>|\\<AL\\>|\\<ALLY\\>|\\<ALEY\\>|\\<ALLE\\>|\\<AY\\>","ALY",x)
  x<-gsub("\\<PLACE\\>|\\<PL\\>|\\<PLC\\>|\\<PLE\\>|\\<PC\\>|\\<PLAC\\>|\\<PLCE\\>|\\<PCE\\>","PL",x)
  x<-gsub("\\<PK\\>|\\<PRK\\>|\\<PRAK\\>|\\<PAK\\>","PARK",x)
  x<-gsub("\\<PARKWAY\\>|\\<PKWY\\>|\\<PARKW\\>|\\<PWY\\>|\\<PKW\\>|\\<PRKWY\\>|\\<PKWY\\>|\\<PKW\\>","PKWY",x)
  x<-gsub("\\<APPR\\>|\\<APR\\>|\\<APPROA\\>|\\<APRCH\\>|\\<APPRCH\\>","APPROACH",x)
  x<-gsub("\\<TERRACE\\>|\\<TER\\>|\\<TERR\\>|\\<TRC\\>|\\<TRCE\\>|\\<TR\\>","TER",x)
  x<-gsub("\\<PLAZA\\>|\\<PLZ\\>|\\<PLAZ\\>|\\<PZ\\>|\\<PLZA\\>","PLZ",x)
  x<-gsub("\\<LANE\\>|\\<LN\\>|\\<LNE\\>|\\<LAN\\>","LN",x)
  x<-gsub("\\<BRIDGE\\>|\\<BRG\\>|\\<BRGD\\>|\\<BGE\\>","BRG",x)
  x<-gsub("\\<HILL\\>|\\<HL\\>|\\<HLL\\>|\\<HIL\\>","HL",x)
  x<-gsub("\\<HEIGHTS\\>|\\<HTS\\>|\\<HT\\>|\\<HEIGHT\\>|\\<HEGHTS\\>|\\<HHT\\>|\\<HEIGT\\>","HTS",x) 
  x<-gsub("\\<SLP\\>|\\<SLEP\\>|\\<SLIIP\\>|\\<SLI\\>","SLIP",x)
  x<-gsub("\\<RON\\>|\\<RW\\>|\\<ROE\\>|\\<ROOW\\>","ROW",x)
  x <- gsub("\\<SQUARE\\>", "SQ", x) # new 
  x<-gsub(".*\\((.*)\\).*", "\\1", x)
  x<-gsub("\\<ST\\>","",x) # new remove streets
  x<-gsub("\\d+\\ - *\\d*|\\d+\\ TO *\\d*|\\d+\\-\\d*","",x) #remove addresses
  
  
  ## dealing with numbered streets
  x<-gsub("(\\d)(ST|ND|RD|TH)\\b", "\\1", x)
  x<-str_remove(x, "(?<=[0-9])(ST|ND|RD|TH)")
  x<-gsub("\\<ONE HUNDRED\\>|\\<ONEHUNDRED\\>|\\<HUNDRED\\>|\\<HUDRED\\>|\\<HUNDED\\>","1",x) 
  x<-gsub("\\<TWO HUNDRED\\>|\\<TWOHUNDRED\\>","2",x)
  x<-gsub("-"," ",x)
  x <- gsub("\\\\", "", x) ##new
  x <- gsub("\\'", "", x) ## remove apostrophe
  x<-gsub("&","AND",x)
  x<-gsub("\\<1ST\\>|\\b1\\b","1",x)
  x<-gsub("\\<2ND\\>|\\b2\\b","2",x)
  x<-gsub("\\<3RD\\>|\\b3\\b","3",x)
  x<-gsub("\\<4TH\\>|\\b4\\b","4",x)
  x<-gsub("\\<5TH\\>|\\b5\\b","5",x)
  x<-gsub("\\<6TH\\>|\\b6\\b","6",x)
  x<-gsub("\\<7TH\\>|\\b7\\b","7",x)
  x<-gsub("\\<8TH\\>|\\b8\\b","8",x)
  x<-gsub("\\<9TH\\>|\\b9\\b","9",x)
  #new
  x<-gsub("\\<TENTH\\>|\\<10TH\\>|\\b10\\b","10",x)
  x<-gsub("\\<ELEVENTH\\>|\\<11TH\\>|\\b11\\b","11",x)
  x<-gsub("\\<TWELFTH\\>|\\<12TH\\>|\\b12\\b","12",x)
  x<-gsub("\\<THIRTEENTH\\>|\\<13TH\\>|\\b13\\b","13",x)
  x<-gsub("\\<FORTEENTH\\>|\\<14TH\\>|\\b14\\b","14",x)
  x<-gsub("\\<FIFTEENTH\\>|\\<15TH\\>|\\b15\\b","15",x)
  x<-gsub("\\<SIXTEENTH\\>|\\<16TH\\>|\\b16\\b","16",x)
  x<-gsub("\\<SEVENTEENTH\\>|\\<17TH\\>|\\b17\\b","17",x)
  x<-gsub("\\<EIGHTEENTH\\>|\\<18TH\\>|\\b18\\b","18",x)
  x<-gsub("\\<NINETEENTH\\>|\\<19TH\\>|\\b19\\b","19",x)
  
  
  x<-gsub("\\<TWENTY\\>|\\<TWENTI\\>|\\<TENTI\\>","2",x)
  x<-gsub("\\<THIRTY\\>|\\<THIRTHY\\>|\\<THIRTEY\\>|\\<TIRTY\\>|\\<TRITHY\\>","3",x)
  x<-gsub("\\<FORTY\\>|\\<FOURTY\\>|\\<FOURTHY\\>|\\<FRTY\\>","4",x)
  x<-gsub("\\<FIFTY\\>|\\<FIFTEY\\>|\\<FIFT\\>|\\<FITY\\>|\\<FIFTHY\\>","5",x)
  x<-gsub("\\<SIXTY\\>|\\<SXTY\\>|\\<SIXY\\>|\\<SXTY\\>|\\<SIXTHY\\>|\\<SIXTEY\\>","6",x)
  x<-gsub("\\<SEVENT\\>|\\<SEVENTY\\>|\\<SEVENTEY\\>|\\<SVENTY\\>|\\<SEVENTI\\>","7",x)
  x<-gsub("\\<EIGHTY\\>|\\<EIGHTEY\\>|\\<EIGHTE\\>","8",x)
  x<-gsub("\\<UNITY\\>|\\<NINTH\\>|\\<NINETY\\>|\\<NINETEY\\>|\\<NINETIETH\\>|\\<NINTY\\>","9",x)
  x<-gsub("\\<FRIST\\>|\\<FIST\\>|\\<FRST\\>|\\<FIRST\\>|\\<ONE\\>","1",x)
  x<-gsub("\\<SECOND\\>|\\<SECORD\\>|\\<SCOND\\>|\\<SECOND\\>|\\<TWO\\>","2",x)
  x<-gsub("\\<THRID\\>|\\<THIRD\\>|\\<TIRD\\>|\\<TRIHD\\>|\\<THREE\\>","3",x)
  x<-gsub("\\<FORTH\\>|\\<FOURTH\\>|\\<FROTH\\>|\\<FROUTH\\>|\\<FOUR\\>","4",x)
  x<-gsub("\\<FIFETH\\>|\\<FIFTH\\>|\\<FIFFTH\\>|\\<FIFTHE\\>|\\<FIVE\\>","5",x)
  x<-gsub("\\<SIXTH\\>|\\<SXTH\\>|\\<SITH\\>|\\<SIHXT\\>|\\<SIX\\>","6",x)
  x<-gsub("\\<SEVENTH\\>|\\<SVEN\\>|\\<SVENTH\\>|\\<SEVENH\\>|\\<SEVENT\\>|\\<SEVEN\\>","7",x)
  x<-gsub("\\<EIGHTH\\>|\\<EIGHTEH\\>|\\<EITH\\>|\\<EIGHT\\>","8",x)
  x<-gsub("\\<NINETH\\>|\\<NINTH\\>|\\<NINT\\>|\\<NINETH\\>|\\<NINE\\>|\\<NIN\\>","9",x)
  x<-gsub("\\<TWENTIETH\\>|\\<TWENTIEFTH\\>","20",x) 
  x<-gsub("\\<THIRTIETH\\>|\\<THIRTIEFTH\\>","30",x)
  x<-gsub("\\<FORTIETH\\>|\\<FOURTIETH\\>","40",x)
  x<-gsub("\\<FIFTIETH\\>","50",x)
  x<-gsub("\\<SIXTIETH\\>","60",x)
  x<-gsub("\\<SEVENTIETH\\>","70",x)
  x<-gsub("\\<EIGHTIETH\\>","80",x)
  x<-gsub("\\<NINETIETH\\>|\\<NINTIETH\\>","90",x)
  x<-gsub("(?<=\\d) (?=\\d)","",x,perl = T) 
  
  x<-gsub("^\\bST\\b","SAINT", x) 
  x<-gsub("\\bHOUSE\\b","", x)
  x<-gsub("\\bHOSTEL\\b","", x)
  x<-gsub("\\bHOTEL\\b","", x)
  x<-gsub("\\bLODGE\\b","", x)
  x<-gsub("\\bLODGING\\b","", x)
  x<-trimws(x, "both")
  
  ## overall spacing problem extra white space
  x <- gsub("((\\w+|\\d+)\\s\\s(\\w+|\\d+))", "\\2 \\3", x)
  x <- gsub("((\\w+|\\d+)\\s\\s\\s(\\w+|\\d+))", "\\2 \\3", x)
  
  ## 4 W to W 4
  x <- gsub("(\\w+|\\d+)\\s(\\bN\\b|\\bW\\b|\\bS\\b|\\bE\\b)", "\\3\\2 \\1", x)
  
  
  ## AVE D
  x <- gsub("(\\b[A-Z])\\s(AVE)", "\\3\\2 \\1", x)

###### ALL MANUAL CLEANING GOES BELOW HERE ######
  
  ## Manual Cleaning
  x <- gsub("\\<ALLANTIC\\>|\\<ATLASTA\\>","ATLANTIC",x)
  x <- gsub("\\<ALLEM\\>","ALLEN",x)
  x <- gsub("\\<ANENUE\\>","AVE",x)
  
  x <- gsub("\\<BATTIE\\>","BALTIC",x)
  x <- gsub("\\<BARREE\\>","BARROW",x)
  x <- gsub("\\<BESSHLEY\\>","BURLING",x)
  x <- gsub("\\<BIRY\\>","BROADWAY",x)
  x <- gsub("\\<BUSTWICK\\>","BUSHWICK",x)
  x <- gsub("\\<BUTTER\\>","BUTLER",x)
  x <- gsub("\\<BREEVORT\\>","BREVOORT",x)
  x <- gsub("\\<BRENNEL\\>|\\<BROOMES\\|\\<BROOM\\|\\<BRANNAS\\>","BROOME",x)

  x <- gsub("\\<CLAIR\\>","CLASSON",x)
  x <- gsub("\\<CLISTEN\\>","CLINTON",x)
  x <- gsub("\\<CHERY\\>","CHERRY",x)
  x <- gsub("\\<CHRYSTEE\\>|\\<CHRYSTAL\\>","CANAL",x)
  x <- gsub("\\<CARRAL\\>|\\<COYAL\\>","CANAL",x)
  x <- gsub("\\<COLUMBIN\\>","COLUMBIA",x)
  x <- gsub("\\<CAMNON\\>","CANNON",x)
  
  x <- gsub("\\<DEVAL\\>","DEVOE",x)
  x <- gsub("\\<DEBOUCHEL\\>","DEBEVOISE",x)
  x <- gsub("\\<DAFONT\\>","DUPONT",x)
  x <- gsub("\\<DEGRAN\\>","DEGRAW",x)
  x <- gsub("\\<DENBO\\>|\\<DEKALB\\>","DE KALB",x)
  x <- gsub("\\<DELAMERE\\>|\\<DALANEY\\>","DELANCEY",x) 

  x <- gsub("\\<ELTHZROTH\\>|\\<ELLSWICK\\>","ELLIOTT",x)
  x <- gsub("\\<ELDREDGE\\>|\\<CLARIDGE\\>","ELDRIDGE",x) 

  x <- gsub("\\<FORSYTHE\\>","FORSYTH",x) 
  x <- gsub("\\<FLATHISH\\>","FLATBUSH",x)

  x <- gsub("\\<GLANCE\\>","GRAND",x) 
  x <- gsub("\\<GOAST\\>","GOERCK",x)
  x <- gsub("\\<GREENS\\>","GREENE",x)
  x <- gsub("\\<GREENRICH\\>|\\<GAMWICH\\>","GREENWICH",x) 
  
  x <- gsub("\\<HAVES\\>","HEWES",x) 
  x <- gsub("\\<HAKEY\\>","HALSEY",x)
  x <- gsub("\\<HEWEY\\>","HENRY",x)
  x <- gsub("\\<HICK\\>","HICKS",x)
  x <- gsub("\\<HUMBOLOT\\>|\\<HUMBARD\\>","HUMBOLDT",x) 
  
  x <- gsub("\\<JOHOM\\>","JOHNS",x) 
  
  x <- gsub("\\<KIOP\\>|\\<HEAP\\>","KEAP",x) 
  
  x <- gsub("\\<LAY FAY ESTE\\>|\\<LADORATT\\>","LAFAYETTE",x)
  x <- gsub("\\<LIRA\\>|\\<LOUMOR\\|\\<LARMER\\>","LORIMER",x)
  x <- gsub("\\<LAAVIUK\\>","LAWRENCE",x) 
  x <- gsub("\\<LAIDLOW\\>","LUDLOW",x) 
  x <- gsub("\\<TEX\\>","LEX",x)
  
  x <- gsub("\\<PARLE\\>|\\<MALLE\\|\\<MYETTE\\>","MYRTLE",x)
  x <- gsub("\\<MC DOUGALL\\>","MCDOUGAL",x)
  x <- gsub("\\<MC DONOUGH\\>","MCDONOUGH",x)
  x <- gsub("\\<MANZA\\>|\\<MAREY\\>","MARCY",x)
  x <- gsub("\\<MESCOLE\\>","MESEROLE",x)
  x <- gsub("\\<MEASE\\>","MOORE",x)
  x <- gsub("\\<MEDDLER\\>","MIDDLETON",x)
  x <- gsub("\\<MANGEN\\>","MANGIN",x)

  x <- gsub("\\<NAPOLK\\>","NORFOLK",x)
  x <- gsub("\\<VAST AND\\>","NOSTRAND",x)

  x <- gsub("\\<OLWEN\\>","OLIVER",x)
  x <- gsub("\\<GERHARD\\>","ORCHARD",x)

  x <- gsub("\\<PUTT\\>","PITT",x)
  x <- gsub("\\<PERROTT\\>","PIERREPONT",x)
  x <- gsub("\\<PLAD\\>","PLACE",x)
  x <- gsub("\\<PRUFER\\>","PROSPECT",x)
  x <- gsub("\\<PREDIDUNT\\>","PRESIDENT",x)
  x <- gsub("\\<PALOKA\\>","PULASKI",x)
     
  x <- gsub("\\<RUTHIE\\>","RUTLEDGE",x)
  x <- gsub("\\<RIDAL\\>","RIDGE",x)
  x <- gsub("\\<REVENTON\\>","RIVINGTON",x)
  x <- gsub("\\<RUALMAINE\\>|\\<RICER SIDE\\>","RIVERSIDE",x)
  
  x <- gsub("\\<SELLTOWN\\>","SULLIVAN",x)
  x <- gsub("\\<SISH\\>","SIDE",x)
  x <- gsub("\\<STUCKER\\>","STEUBEN",x)
  x <- gsub("\\<SUMME\\>","SUMMIT",x)
  x <- gsub("\\<SCHOMERDOSA\\>","SCHERMERHORN",x)
  x <- gsub("\\<DOUTH\\>|\\<SONSE\\>","SOUTH",x)
  
   x <- gsub("\\<STONPSON\\>","THOMPSON",x)
  x <- gsub("\\<TRAY\\>","TROY",x)
  x <- gsub("\\<TAYLER\\>","TAYLOR",x)
  
  x <- gsub("\\<WMON\\>","UNION",x)
  
  x <- gsub("\\<WAR CAREN\\>","VAN BUREN",x)
  x <- gsub("\\<VEMON\\>","VERNON",x)
  x <- gsub("\\<VANDERLY\\>|\\<VANDERSLIDE\\>","VANDERBILT",x)
  
  x <- gsub("\\<WITKINS\\>","WATKINS",x)
  x <- gsub("\\<WALLWORTH\\>","WALWORTH",x)
  x <- gsub("\\<WHIPPER\\>","WHIPPLE",x)
  x <- gsub("\\<WALLABANK\\>","WALLABOUT",x)
  x <- gsub("\\<WASH\\>|\\<WASTEWATER\\>","WASHINGTON",x)
    
}  
```

#### Street Matching Function `02_Street_Matching_MNBK` (STREET CLEAN FIRST)
```{r}
# ---- CLEAN ADDRESSES ----
## Extract Unique Addresses
unique_addresses <- sample %>%
    select(ED, street_add) %>%
    filter(!is.na(street_add)) %>%
    distinct(ED, street_add)

  ## import rem_dup_word
  rem_dup_word <- function(x){
    x <- tolower(x)
    paste(unique(trimws(unlist(strsplit(x,split=" ",fixed=F,perl=T)))),collapse = 
            " ")
  }
  
  ## [NOTE] make sure to run clean() function from '01_Street_Clean_Function_MNBK.R'
  
  ## preallocate memory for cleaned column
  cleaned <- rep(NA_character_, nrow(unique_addresses))
  
  ## apply rem_dup_word() and str_clean()
  for (i in 1:nrow(unique_addresses)) {
    cleaned[i] <- toupper(rem_dup_word(unique_addresses[i, 2]))
    cleaned[i] <- clean(cleaned[i])
  }
```

#### Street Matching Function `02_Street_Matching_MNBK` (FULL RUN)
```{r street match}
street_match <- function(sample, edict) {

  ## clean ED number
  edict <- edict %>%
    mutate(ED = str_pad(ED, 4, "left", pad = "0")) # ensure numbers are all 4 digit
  
  # ---- FIRST MATCH ----
  ## create add_matches, which combines `unique_addresses` with the `cleaned` column
  add_matches <- tibble(ED = unique_addresses$ED,
                        raw = unique_addresses$street_add,
                        clean = cleaned)
  
  
  ## create str_algo, which has all the methods we will use
  str_algo <- c("dl", "qgram", "cosine", "jaccard", "jw")
  
  ## create empty lists which will store outputs
  algo_list <- list()
  ed_list <- list()
  
  ## matching
  for (i in unique(add_matches$ED)) {
    # create subset of edict and add_matches for a particular ED
    # TO ACCEPT BK: ED_dict <- filter(edict, ED == i & MNBK == "BK") %>%
    ED_dict <- filter(edict, ED == i) %>% 
      select(- "ED") %>%
      unlist() 
    ED_dict <- data.frame(clean = ED_dict, stringsAsFactors = FALSE) %>%
      filter(!is.na(clean))
    ED_add <- filter(add_matches, ED == i)
    
    for (j in str_algo) {
      # stringdist_join, extract best match for each method
      result <- stringdist_left_join(ED_add, ED_dict, by = "clean", 
                                     max_dist = 5, method = j, distance_col = "dscore") %>%
        select(ED, raw, clean.x, clean.y, dscore) %>%
        group_by(raw) %>%
        arrange(dscore) %>%
        slice(1)
      
      # at this point, result = dataframe of addresses in ED (i), merged with method (j)
      # append result to algo_list, each element in algo_list is a dataframe
      algo_list[[j]] <- result
      
      # "column bind" all elements in algo_list to form a df called ed_df
      # ed_df now contains all addresses in ED(i) with matches using all methods
      ed_df <- algo_list %>% 
        reduce(left_join, by = c("ED", "raw", "clean.x"))
    }
    # append each ed_df to ed_list 
    ed_list[[i]] <- ed_df
  }
  
  ## bind all dfs by row
  match_dict <- bind_rows(ed_list) %>%
    select(ED = "ED", raw = "raw", clean = "clean.x",
           match_dl = "clean.y.x", dscore_dl = "dscore.x",
           match_qgram = "clean.y.y", dscore_qgram = "dscore.y",
           match_cos = "clean.y.x.x", dscore_cos = "dscore.x.x",
           match_jac = "clean.y.y.y", dscore_jac = "dscore.y.y",
           match_jw = "clean.y", dscore_jw = "dscore")
  
  ## ---- FURTHER TUNING ----
  ## create summary statistics
  summ_dscores <- match_dict %>%
    replace(is.na(.), 5) %>%
    group_by(ED) %>%
    summarize(mean_dl = mean(dscore_dl), sd_dl = sd(dscore_dl),
              mean_qgram = mean(dscore_qgram), sd_qgram = sd(dscore_qgram),
              mean_cos = mean(dscore_cos), sd_cos = sd(dscore_cos),
              mean_jac = mean(dscore_jac), sd_jac = sd(dscore_jac),
              mean_jw = mean(dscore_jw), sd_jw = sd(dscore_jw)) %>%
    replace(is.na(.), 0)
  
  ## format into a list such that e.g. `summ_dscores_list$0010$jw$mean` gives mean jw dscore of ED 0010
  summ_dscores_list <- unique(summ_dscores$ED) %>% as.list()
  names(summ_dscores_list) <- unique(summ_dscores$ED)
  
  for (i in names(summ_dscores_list)) {
    s <- filter(summ_dscores, ED == i)
    ED_sum_list <- list(dl = list(mean = s$mean_dl, sd = s$sd_dl),
                        qgram = list(mean = s$mean_qgram, sd = s$sd_qgram),
                        cosine = list(mean = s$mean_cos, sd = s$sd_cos),
                        jaccard = list(mean = s$mean_jac, sd = s$sd_jac),
                        jw = list(mean = s$mean_jw, sd = s$sd_jw))
    summ_dscores_list[[i]] <- ED_sum_list
  }
  
  ## match again
  for (i in unique(add_matches$ED)) {
    # create subset of edict and add_matches for a particular ED
    # TO ACCEPT BK: ED_dict <- filter(edict, ED == i & MNBK == "BK") %>%
    ED_dict <- filter(edict, ED == i) %>% 
      select(- "ED") %>%
      unlist() 
    ED_dict <- data.frame(clean = ED_dict, stringsAsFactors = FALSE) %>%
      filter(!is.na(clean))
    ED_add <- filter(add_matches, ED == i)
    
    for (j in str_algo) {
      # set a max_dist that varies by ED. max_dist is 2sd higher than mean.
      threshold <- summ_dscores_list[[i]][[j]]$mean + 2 * summ_dscores_list[[i]][[j]]$sd
      
      # stringdist_join, extract best match for each method
      result <- stringdist_left_join(ED_add, ED_dict, by = "clean", 
                                     max_dist = threshold, method = j, distance_col = "dscore") %>%
        select(ED, raw, clean.x, clean.y, dscore) %>%
        group_by(raw) %>%
        arrange(dscore) %>%
        slice(1)
      
      # at this point, result = dataframe of addresses in ED (i), merged with method (j)
      # append result to algo_list, each element in algo_list is a dataframe
      algo_list[[j]] <- result
      
      # "column bind" all elements in algo_list to form a df called ed_df
      # ed_df now contains all addresses in ED(i) with matches using all methods
      ed_df <- algo_list %>% 
        reduce(left_join, by = c("ED", "raw", "clean.x"))
    }
    # append each ed_df to ed_list 
    ed_list[[i]] <- ed_df
  }
  
  ## bind all dfs by row
  match_dict <- bind_rows(ed_list) %>%
    select(ED = "ED", raw = "raw", clean = "clean.x",
           match_dl = "clean.y.x", dscore_dl = "dscore.x",
           match_qgram = "clean.y.y", dscore_qgram = "dscore.y",
           match_cos = "clean.y.x.x", dscore_cos = "dscore.x.x",
           match_jac = "clean.y.y.y", dscore_jac = "dscore.y.y",
           match_jw = "clean.y", dscore_jw = "dscore")
  
  # ---- BEST MATCH ----
  ## function to extract modes
  modal <- function(x) {
    freq <- table(x) %>% as.data.frame(stringsAsFactors = FALSE)
    result <- freq[which(freq$Freq == max(freq$Freq)), 1]
    if (length(result) > 1){
      result <- str_c(result, collapse = "+")
    }
    result
  }
  
  ## extract modes
  match_dict_modes <- match_dict %>%
    replace(is.na(.), "0") %>%
    select(- c("dscore_dl", "dscore_qgram", "dscore_cos", "dscore_jac", "dscore_jw")) %>%
    gather("method", "match", - c("ED", "raw", "clean")) %>%
    group_by(ED, raw) %>%
    summarise(mode = modal(match))
  
  ## join to match_dict
  match_dict <- left_join(match_dict, match_dict_modes, by = c("ED", "raw"))
  
  ## best match
  match_dict <- match_dict %>%
    replace(is.na(.), 0) %>%
    rowwise() %>%
    mutate(result_type = ifelse(match_dl == match_cos & match_dl == match_qgram & match_dl == match_jac & match_dl == match_jw, ifelse(clean == match_dl, 1, ifelse(mode == 0, 6, 2)), ifelse(str_detect(mode, "\\+"), 4, ifelse(mode == 0, 5, 3)))) %>%
    mutate(best_match = case_when(result_type %in% c(1, 2, 3) ~ mode,
                                  result_type %in% c(4, 5, 6) ~ "0"))
  
  # ---- FILL DOWN ----
  ## create flags for row that had to be filled
  match_dict <- mutate(match_dict, flag = ifelse(best_match == 0, 1, 0))
  match_dict$index <- 1:nrow(match_dict)
  
  na_matches <- filter(match_dict, best_match == 0)
  
  ## fill down
  for (i in na_matches$index) {
    # extract pool of potential matches
    i <- as.numeric(i)
    match_ED <- na_matches[na_matches$index == i, "ED"]
    pool_indexes <- c(i - 3, i - 2, i - 1, i + 1, i + 2, i + 3)
    pool_matches <- filter(match_dict, index %in% pool_indexes & ED == match_ED) %>%
      select(best_match) %>%
      distinct()
    
    # match using jw
    match <- stringdist_left_join(filter(na_matches, index == i), pool_matches, by = c(clean = "best_match"), 
                                  max_dist = 5, method = "jw", distance_col = "dscore") %>%
      group_by(raw) %>%
      arrange(dscore) %>%
      slice(1)
    
    # sub match into match_dict
    match_dict[match_dict$index == i, "best_match"] <- match$best_match.y
  }
  
  # ---- MERGE TO ORIGINAL SAMPLE ----
  match_dict_subset <- select(match_dict, ED, raw, best_match, result_type, flag_st = flag)
  sample_cleaned <- left_join(sample, match_dict_subset, by = c("ED" = "ED", "street_add" = "raw"))
}
```

#### Match Street Fill Down Function `03_Matched_Street_Fill_Down`
```{r match street fill down}
fillDownStreet <- function(df){
  
  df <- df %>%
    filter(record == "H") %>%
    mutate(best_match_temp = best_match) %>% 
    group_by(ED) %>%
    fill(street_add, best_match, result_type, .direction="down") %>%
    fill(street_add, best_match, result_type, .direction = "up") %>%
    rowwise() %>% 
    mutate(flg_filled_st = ifelse(!is.na(best_match)  && is.na(best_match_temp), 1, 0)) %>% 
    select(-best_match_temp) %>% 
    ungroup()
  
  return(df)
}
```

#### Clean House Number Function `04_house_clean`
```{r clean house number}
house_clean <- function(sample, edict, hn_dict) {
  
  # ---- Preprocessing ----
  # select only records with house numbers

  HN <- sample %>%
    fill(microfilm, .direction = "up") %>%
    filter(!is.na(house_num)) %>%
    select(microfilm, ED, 
           dwelling_ser, house_num, street_add, best_match) %>%
    mutate(originalHN = house_num,
           house_num = ifelse(house_num=="", NA, as.character(house_num)),
           street_add = ifelse(is.na(street_add), " ", street_add))
  
  # ---- Cleaning: Extract House No. + Modifiers ----
  # Extract house number from street
  HN <- HN %>% rowwise() %>%
    mutate(flag1=ifelse(is.na(house_num)&
                          !is.na(str_extract(street_add,"[0-9]+\\s*(-|TO)+\\s*[0-9]+")), 
                        str_extract(street_add,"[0-9]+\\s*(-|TO)+\\s*[0-9]+"),
                        NA),
           house_num=ifelse(is.na(house_num), flag1, house_num),
           flg_hn_from_strt2=ifelse((!is.na(flag1))&(house_num != flag1) , 1, 0),
           flg_hn_from_strt=ifelse(is.na(flag1), 0, 1),
           house_num = gsub("\\s*(TO)\\s*", "-", house_num, ignore.case = TRUE)) %>%
    select(-flag1)
  
  # Separate modifiers into new column
  HN <- HN %>% rowwise() %>%
    mutate(modifier.number = str_extract(house_num, "\\s1[/-]2\\b"), 
           house_num = gsub("\\s1[/-]2\\b", " ", house_num),  
           house_num = gsub("\\s*(TO|/|&)\\s*", "-", house_num, ignore.case = TRUE), 
           house_num = gsub("(?<=\\d)\\s+(?=\\d)", "-", house_num, ignore.case = TRUE, perl = TRUE), 
           modifier.word = trimws(str_extract(house_num, "[A-Za-z\\s]+")),            
           house_num = gsub("\\s+", " ", house_num),                    
           house_num = trimws(gsub("[A-Za-z\\s]+", "", house_num)),
           house_num = gsub("^\\D+", "", house_num, ignore.case = TRUE), 
           house_num = gsub("\\D+$", "", house_num, ignore.case = TRUE), 
           flg_cleaned = ifelse(originalHN=="", 0, ifelse(house_num==originalHN, 0, 1))) 
  
  # ---- Cleaning: House Ranges ----
  # Cleaning house ranges with "-"
  hn_range_clean <- function(hn_range) {
    hn <- hn_range %>%
      str_split("-") %>% unlist()
    
    same_length <- str_length(hn) %>%
      unique() %>%
      length() == 1
    
    if (!same_length) {
      hn <- as.integer(hn)
      max_hn <- as.character(max(hn))
      min_hn <- as.character(min(hn))
      len_diff <- str_length(max_hn) - str_length(min_hn)
      
      clean <- map_chr(hn, function(x) ifelse(str_length(x) < str_length(max_hn), 
                                              ifelse(str_length(paste0(str_sub(max_hn, 1, len_diff), x)) == str_length(max_hn),
                                                     paste0(str_sub(max_hn, 1, len_diff), x),
                                                     x),
                                              x)
      )
      
      hn_range <- str_c(clean, collapse = "-")
      
    }
    
    return(hn_range)
  }
  
  HN <- HN %>%
    mutate(hn_range = hn_range_clean(house_num)) %>%
    mutate(flg_cleaned = ifelse(house_num == hn_range, 0, 1)) 
  
  splt_df <- str_split_fixed(HN$hn_range, pattern = "-", n = 3) %>% data.frame() %>% 
    rename(hn_1 = X1, hn_2 = X2, hn_3 = X3) %>% 
    mutate_all(as.character) %>% mutate_all(as.numeric)
  
  HN <- HN  %>% 
    cbind(splt_df)
  
  # Additional cleaning of large house numbers that are ranges
  HN <- HN %>%
    left_join(hn_dict, by = c("ED" = "ED", "best_match" = "Name"))
  
  hn_ranges <- HN %>%
    filter(str_length(hn_1) >= 4) %>%
    filter(str_length(hn_1) != str_length(Low) & str_length(hn_1) != str_length(High)) %>%
    mutate(hn1_first_2 = as.integer(str_sub(hn_1, 1, 2))) %>%
    mutate(hn1_last_2 = as.integer(str_sub(hn_1, str_length(hn_1) - 1, str_length(hn_1)))) %>%
    filter(hn1_first_2 >= Low & hn1_first_2 <= High & hn1_last_2 >= Low & hn1_last_2 <= High) %>%
    distinct(house_num) %>% unlist()

  HN <- HN %>%
    mutate(hn1_first_2 = as.integer(str_sub(house_num, 1, 2)),
           hn1_last_2 = as.integer(str_sub(house_num, str_length(house_num) - 1, str_length(house_num)))) %>%
    mutate(hn_1 = ifelse(house_num %in% hn_ranges, hn1_first_2, hn_1),
           hn_2 = ifelse(house_num %in% hn_ranges, hn1_last_2, hn_2)) %>%
    mutate(flg_cleaned = ifelse(originalHN != hn_1, 1, 0))
  
  # ---- Output ----
  # [FLAG]: choose variables
  HN <- HN %>%
    select(ED, dwelling_ser, originalHN, best_match, 
           modifier.number, modifier.word,
           hn_1, hn_2, hn_3, flag_hn_cleaned = flg_cleaned) %>%
    distinct(ED, dwelling_ser, originalHN, best_match, .keep_all = TRUE)
  
  sample <- left_join(sample, HN,
                      by = c("ED", "dwelling_ser", "best_match",
                             "house_num" = "originalHN"))
  
  return(sample)
}
```

#### House Number Fill Down `05_House_Number_Fill_Down`
```{r house number fill down}
fillDownHouseNum <- function(df){
  
  df <- df %>% 
    filter(record == "H") %>%
    mutate(house_num_temp = house_num) %>% 
    group_by(ED, best_match) %>%
    fill(modifier.number, modifier.word, house_num, hn_1, hn_2, hn_3, .direction="down") %>%
    fill(modifier.number, modifier.word, house_num, hn_1, hn_2, hn_3, .direction="up") %>%
    rowwise() %>% 
    mutate(flg_filled_hn = ifelse(!is.na(house_num)  && is.na(house_num_temp), 1, 0)) %>% 
    select(- house_num_temp) %>%
    ungroup()
  return(df)
}
```

<br/>

### Run Scripts
***
```{r, cache=TRUE}
output <- sample %>%
  street_match(., edict) %>% # from 02_Street_Matching
  fillDownStreet(.) %>% # from 03_Matched_Street_Fill_Down
  house_clean(., edict, hn_dict) %>% # from 04_house_clean
  fillDownHouseNum(.) # from 05_House_Number_Fill_Down
```

### Save Output Folder
```{r}
#output
#write.csv(output, file = "mn_output.csv")
#saveRDS(output, file = "mn_output.rds")
```
