---
title: "MN_1910"
author: "Clinton"
date: "21 May 2020"
output: html_document
---


```{r message = FALSE, warning = FALSE, echo = FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(DT)
library(sf)

output = read_csv("output.csv",
                  col_types = cols(modifier.number = col_integer()))

theme_set(theme_minimal())
```

### Graph of output from the Manhattan cleaning script

```{r echo = FALSE}
output %>%
  group_by(result_type) %>%
  summarise(count = n()) %>%
  mutate(countP = round(count / NROW(output) * 100, digits = 0),
         label = paste0(count, " (", countP, "%)")) %>%
  ggplot(aes(y = count, x = factor(result_type))) +
  geom_col(fill = "#ffa600") +
  geom_text(aes(label = label), 
            vjust = -0.5,
            size = 3) +
  
  scale_y_continuous(limits = c(0, 220000),
                     name = "Count") +
  labs(x = "Result Type")
```

<br>

In comparison, the previous results were:

* 34.7% Perfect Match
* 25.4% Identical Match
* 29.8% Singular Mode
* 2% Multiple Modes
* 0.8% NAs
* 0.3% No Match

<br>

### Problematic results

#### Table of results which were matched as 5 or 6

```{r echo = FALSE}
output %>%
  filter(result_type == 5 | result_type == 6) %>% 
  select(ED, street_add, best_match, result_type) %>% 
  arrange(ED) %>% 
  dplyr::distinct() %>%
  datatable()
```

<br>

#### Table of problematic EDs

```{r echo = FALSE}
output %>%
  group_by(ED) %>%
  mutate(size = n()) %>%
  filter(result_type == 5 | result_type == 6) %>% 
  group_by(ED, size, result_type) %>%
  summarise(count = n()) %>%
  mutate(count_prop = round(count / size * 100, digits = 2)) %>%
  select(ED, result_type, count, count_prop) %>%
  arrange(desc(count)) %>%
  datatable()
```

<br>

The trend is similar to previous findings, that EDs 0815, 1481 and 0009 have a large number of mismatches

<br>

#### Graph of mismatches

```{r echo = FALSE, message = FALSE, warning = FALSE}
output %>%
  group_by(ED) %>%
  summarise(mean_result_type = mean(result_type),
            size = n()) %>%
  mutate(ED = as.numeric(ED)) %>%
  
  ggplot(aes(y = mean_result_type, 
             x = ED, 
             col = size)) +
  geom_point(alpha = 0.9) + 
  labs(y = "Average result type") +
  scale_colour_viridis_c() + 
  theme(legend.position = "top",
        legend.key.width = unit(3,"line"))
```

<br>

#### Map of Average Result by ED

```{r echo = FALSE, messag= FALSE}
mn_1910 = st_read("shp/Manhattan_1910.shp",
                  stringsAsFactors = FALSE)

output %>%
  group_by(ED) %>%
  summarise(avg_result = mean(result_type)) %>%
  mutate(ED = as.numeric(ED),
         ED = as.character(ED)) %>%
  left_join(mn_1910, ., by = "ED") %>%
  
  ggplot() +
  geom_sf(aes(fill = avg_result),
          colour = "black",
          size = 0.25) +
  scale_fill_viridis_c() +
  labs(fill = "Average result")
```


