---
title: "House No. Cleaning"
author: "Jolene Lim"
date: "8 December 2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

# 1. Load Data
```{r, message = FALSE}
sample <- read_csv("data/us1910m_usa_sample100k.csv")

names(sample) <- c("hnyc_id", "record", "township", "county", "ED", "person_id",
              "dwelling_seq", "dwelling_ser", "dwelling_ser2", "hh_seq", "hh_seq_8",
              "hh_ser2", "hh_ser_bef_split", "indiv_seq", "split",
              "line_no", "line_no2", "microfilm",
              "n_fam", "n_person_bef_split", "house_num", "street_add")
```

# 2. Separate House Numbers
Note: code is the same as Bo's `in_progress_HN_html` script, look at that script for more documentation
Preprocessing to filter household persons
```{r}
HN <- sample %>%
    fill(microfilm, .direction = "up") %>%
    filter(record=="H") %>%
    select(microfilm, ED, 
           dwelling_ser, house_num, street_add) %>%
    mutate(originalHN = house_num,
           house_num = ifelse(house_num=="", NA, as.character(house_num)),
           street_add = ifelse(is.na(street_add), " ", street_add))
```

Extract house number
```{r}
HN <- HN %>% rowwise() %>%
  mutate(flag1=ifelse(is.na(house_num)&
                               !is.na(str_extract(street_add,"[0-9]+\\s*(-|TO)+\\s*[0-9]+")), 
                             str_extract(street_add,"[0-9]+\\s*(-|TO)+\\s*[0-9]+"),
                             NA),
         house_num=ifelse(is.na(house_num), flag1, house_num),
         flg_hn_from_strt2=ifelse((!is.na(flag1))&(house_num != flag1) , 1, 0),
         flg_hn_from_strt=ifelse(is.na(flag1), 0, 1),
         house_num = gsub("\\s*(TO)\\s*", "-", house_num, ignore.case = TRUE)) %>%
  select(-flag1)
```

Cleaning
The chunk below does:

1. Copy `1/2` or `1-2` from `house_num` to `modifier.number` column
2. Remove `1/2` or `1-2` from `house_num`
3. Change `TO` and `/` into `-`
4. Copy any modifier words from `house_num` to `modifier.number` column
5. Remove excess whitespaces
6. Remove any modifier words from `house_num`

** I include `1-2` just in case this exists. The regex for this case excludes `11-20` for example.
```{r}
HN <- HN %>% rowwise() %>%
  mutate(modifier.number = str_extract(house_num, "\\s1[/-]2\\b"), 
         house_num = gsub("\\s1[/-]2\\b", " ", house_num),  
         house_num = gsub("\\s*(TO|/|&)\\s*", "-", house_num, ignore.case = TRUE), 
         house_num = gsub("(?<=\\d)\\s+(?=\\d)", "-", house_num, ignore.case = TRUE, perl = TRUE), 
         modifier.word = trimws(str_extract(house_num, "[A-Za-z\\s]+")),            
         house_num = gsub("\\s+", " ", house_num),                    
         house_num = trimws(gsub("[A-Za-z\\s]+", "", house_num)),
         house_num = gsub("^\\D+", "", house_num, ignore.case = TRUE), 
         house_num = gsub("\\D+$", "", house_num, ignore.case = TRUE), 
         flg_cleaned = ifelse(originalHN=="", 0, ifelse(house_num==originalHN, 0, 1))) 
```

Splitting house ranges
```{r}
splt_df <- str_split_fixed(HN$house_num, "-", 3) %>% data.frame() %>% rename(hn_1 = X1, hn_2 = X2, hn_3 = X3) %>% mutate_all(as.character) %>% mutate_all(as.numeric)
HN <- HN  %>% cbind(splt_df) 
```

# 3. Fill Down
Basic Principle: if the first non-NA above and first non-NA below has a difference of 0 or 2, AND the ED of those are the same, AND the street address are the same, fill down.

```{r}
approach_1 <- HN %>%
  
  # fix street_add
  mutate(street_add = as.character(street_add)) %>%
  
  # create copy of the columns to filled up/down - these should only be filled for records with house no.s
  mutate(above_hn = hn_1, below_hn = hn_1) %>% 
  mutate(above_ED = ifelse(!is.na(hn_1), ED, NA), below_ED = ifelse(!is.na(hn_1), ED, NA)) %>%
  mutate(above_add = ifelse(!is.na(hn_1), street_add, NA), 
         below_add = ifelse(!is.na(hn_1), street_add, NA)) %>%
  
  # fill
  fill(above_hn, .direction = "down") %>%
  fill(below_hn, .direction = "up") %>%
  fill(above_ED, .direction = "down") %>%
  fill(below_ED, .direction = "up") %>%
  fill(above_add, .direction = "down") %>%
  fill(below_add, .direction = "up") %>%
  
  # can we confidently fill up/down?
  mutate(yes_fill = ifelse(abs(above_hn - below_hn) %in% c(0, 2)
                           & above_ED == below_ED & above_ED == ED
                           & above_add == below_add & above_add == street_add
                           & is.na(hn_2), 
                           1, 0))

# how many confident fill downs?
sum(approach_1$yes_fill)
sum(approach_1$yes_fill) - sum(!is.na(approach_1$hn_1))
```

Using this method, 5806 additional records have been recovered (with confidence), i.e. we have 10660 confident house numbers. This number is likely to increase with cleaned street addresses.

Code to actually do the fill down (above is just EDA). Chunk set to `eval = FALSE` for now. Run this code instead of chunk above.
```{r, eval = FALSE}
HN2 <- HN %>%
  
  # fix street_add
  mutate(street_add = as.character(street_add)) %>%
  
  # create copy of the columns to filled up/down - these should only be filled for records with house no.s
  mutate(above_hn = hn_1, below_hn = hn_1) %>% 
  mutate(above_ED = ifelse(!is.na(hn_1), ED, NA), below_ED = ifelse(!is.na(hn_1), ED, NA)) %>%
  mutate(above_add = ifelse(!is.na(hn_1), street_add, NA), 
         below_add = ifelse(!is.na(hn_1), street_add, NA)) %>%
  
  # fill
  fill(above_hn, .direction = "down") %>%
  fill(below_hn, .direction = "up") %>%
  fill(above_ED, .direction = "down") %>%
  fill(below_ED, .direction = "up") %>%
  fill(above_add, .direction = "down") %>%
  fill(below_add, .direction = "up") %>%
  
  # replace hn_1 column
  mutate(hn_1 = ifelse(abs(above_hn - below_hn) %in% c(0, 2) 
                       & above_ED == below_ED & above_ED == ED
                       & above_add == below_add & above_add == street_add
                       & is.na(hn_2), 
                       above_hn, NA_integer_))
```

# 4. Overly Large House Numbers
Whether a house number is considered large is dependent on the numbers above and below, e.g. if 9092 (which would actually be 90-92) is surrounded by 86 and 94, then clearly this is a large house number.

```{r}
large_hn <- HN2 %>%
  mutate(hn1_first_2 = as.integer(str_sub(hn_1, 1, 2))) %>%
  mutate(hn1_last_2 = as.integer(str_sub(hn_1, str_length(hn_1) - 1, str_length(hn_1))))

large_hn$next_hn <- large_hn$hn_1 %>% unlist() %>%
  append(NA_integer_, after = length(.)) %>%
  .[-1]
  
large_hn$prev_hn <- large_hn$hn_1 %>% unlist() %>%
  append(NA_integer_, after = 0) %>%
  .[-length(.)]

large_hn$prev_street_add <- large_hn$street_add %>% unlist() %>%
  append(NA_character_, after = 0) %>%
  .[-length(.)]

large_hn$next_street_add <- large_hn$street_add %>% unlist() %>%
  append(NA_character_, after = length(.)) %>%
  .[-length(.)]

large_hn %>%
  filter(str_length(hn_1) != str_length(next_hn) 
         & str_length(hn_1) != str_length(prev_hn)
         & street_add == prev_street_add
         & street_add == next_street_add) %>%
  select(ED, house_num, hn_1, hn1_first_2, hn1_last_2, prev_hn, next_hn) %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px")

# to only select potential house number ranges
## not implemented; this was the original prupose of the EDA but EDA found more issues, but good to have a reference code for selecting specifically house number ranges
#large_hn %>%
#  filter(str_length(hn_1) != str_length(next_hn) & str_length(hn_1) != str_length(prev_hn)) %>%
#  filter(hn1_first_2 - hn1_last_2 < 10) %>%
#  filter(str_length(hn_1) >= 4) %>%
#  select(ED, house_num, hn_1, hn1_first_2, hn1_last_2, prev_hn, next_hn) %>%
#  kable() %>% kable_styling()
```