---
title: "linear_filldown"
author: "Bo Jumrustanasan"
date: "12/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Approach not fully explored. May be dropped. 

## Linear Fill-in

```{r}
## evenlyFill.R
<<evenly_fill_HN7>>
```

merge_SEQ 2 and 3 (top) and 501 and 502 (bottom). Blue dots are real house numbers. Red dots are interpolations.

```{r}
## need filled_df from evenlyFill.R
## plot house number movement by merged sequences
filled_hn <- filled_df %>% select(i, house_num, hn_filled, merge_SEQ) %>%
  mutate(hn_filled = ifelse(!is.na(house_num), NA, hn_filled),
         merge_SEQ = as.factor(merge_SEQ)) %>%
  gather(key = hn, value = value, -i, -merge_SEQ)

mseq_list <- filled_df %>% pull(merge_SEQ) %>% unique()
plt1 <- ggplot(filled_hn %>% filter(merge_SEQ %in% mseq_list[c(2,3)]), 
               aes(x=i, y = value, color = hn)) + 
  geom_point(aes(shape = merge_SEQ)) + 
  theme(legend.position = "none")
plt2 <- ggplot(filled_hn %>% filter(merge_SEQ %in% mseq_list[c(501, 502)]), 
               aes(x=i, y = value, color = hn)) + 
  geom_point(aes(shape = merge_SEQ)) + 
  theme(legend.position = "none")
grid.arrange(plt1, plt2, ncol = 1)
```

Missing house numbers before the filled in.

```{r}
filled_df %>% select(house_num) %>% 
  mutate(na_cnt = ifelse(is.na(house_num), 1, 0)) %>% 
  summarize(missing = sum(na_cnt), ttl = n()) %>% kable %>% kable_styling()
```

Missing house numbers after the filled in. Missing values are those at the end of sequences.

```{r}
filled_df %>% select(hn_filled) %>% 
  mutate(na_cnt = ifelse(is.na(hn_filled), 1, 0)) %>% 
  summarize(missing = sum(na_cnt), ttl = n()) %>% kable %>% kable_styling()
```

Records with filled in house numbers are flagged in column `flg_hn_filled`. Records with SEQ==`NA` will have `flg_hn_filled`==`NA`.

```{r}
filled_df %>% head(100) %>% kable %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px")
```


