---
title: "House Number Sequence"
author: "Bo Jumrustanasan"
date: "12/10/19"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(gridExtra)
```

The long string of house numbers from enumeration pages is divided into segments, referred to as **sequences**. A sequence is expected to contain 1 street with reasonable running house numbers. The process is for that missing house numbers and street names within a sequence can be confidently recovered.

## Sequence Detection

There are two main steps to detect sequences: breaking and merging. They are wrapped under a function called `getMergedSeq()` which return a dataframe with a column of sequence numbers.

### Breaking Step

The string of house number is first divided into small segments, referred to as **subsequences**, so that within each subsequence, 

1. there is only one parity in house numbers
2. house number is only increasing (decreasing)
3. there is only one street name (only when `check_street` is set to TRUE)

See `getSequence.R` for the script.


### Merging Step

In reality, an enumerator could go back and forth from house to house with in the same street. Thus, there could be both increasing and decreasing house numbers in the same sequences. Moreover, having only one parity can be too restrictive as a number can be mistranscribed into another that has different parity. 
ƒ
This steps merges back some subsequences into sequences so that, within a sequence,

1. there is only 1 street name (only when `check_street` is set to TRUE)
2. there are no more than 2 house numbers whose parities are different from the rest of the house numbers in the sequence
3. Adjacent house numbers to not differ for more than 10. The number can be set by setting `jump_size`.

See `mergeSequence.R` for the script.

```{r}
#load("Scripts/House/HN5.RData")
load("HN5.RData")
```

```{r}
## Tell R we will read in codes from these files.
read_chunk("getSequence.R")
read_chunk("mergeSequence.R")
read_chunk("evenlyFill.R")
read_chunk("lookAroundBP.R")
```
    
```{r}
#' Read in functions from files into Rmd

## getSequece.R
<<getSeq_fxs>>
  
## mergeSequence.R
<<read_merge_fxs>>
<<create_HN7>>
  
## lookAroundBP.R
<<lookAroundBP>>
```

```{r}
#' Process dataframes for EDA
HN7_hn <- getMergedSeq(HN5, jump_size = 500, check_street = FALSE) # Ignore street names
HN7_st <- getMergedSeq(HN5, jump_size = 500, check_street = TRUE) # Consider street names
```

## House numbers by `SEQ` and `merged_SEQ`

**On us1910m_usa_sample100k.csv**

### Sample census data with `check_street` off

Sample rows with their subsequence number and merged sequence number.

* 18 Madison is treated as a mistranscription because it is the only even number among its neighboring subsequences. **very crucial to fix this before filling in missing house numbers**
* 3 Pearl is merged to the previous sequence. Moving to the new street has not been included as a condition to merge.
* 430 Pearl is not merged to the previous subsequence because the house number jump (from 3 to 430) is too large.

See below.

```{r}
x <- HN7_hn %>% select(house_num, street_add, SEQ, merge_SEQ)
x[200:500,] %>% kable %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px")
```


### Sample census data with `check_street` on

```{r}
x <- HN7_st %>% select(house_num, street_add, SEQ, merge_SEQ)
x[200:500,] %>% kable %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px")
```

### Overview

Left: house number *subsequences*

Middle: house number *merged sequences* + street off

Right: house number *merged sequences* + street on

```{r, fig.width=15}
sub_seq <- ggplot(HN7_hn[700:1000,], aes(x = i, y = house_num, color = SEQ)) + 
  geom_point() +
  xlab("Record Index") + 
  ylab("House Number") + 
  ggtitle("Sample 1: Subsequences")

seq <- ggplot(HN7_hn[700:1000,], aes(x = i, y = house_num, color = merge_SEQ)) + 
  geom_point() +
  xlab("Record Index") + 
  ylab("House Number") + 
  ggtitle("Sample 1: Merged Sequences (street off)")

seq_st <- ggplot(HN7_st[700:1000,], aes(x = i, y = house_num, color = merge_SEQ)) + 
  geom_point() +
  xlab("Record Index") + 
  ylab("House Number") + 
  ggtitle("Sample 1: Merged Sequences (street on)")

grid.arrange(sub_seq, seq, seq_st,  ncol = 3)
```


```{r, fig.width=15}
sub_seq <- ggplot(HN7_hn[3100:3400,], aes(x = i, y = house_num, color = SEQ)) + 
  geom_point() +
  xlab("Record Index") + 
  ylab("House Number") + 
  ggtitle("Sample 2: Subsequences")

seq <- ggplot(HN7_hn[3100:3400,], aes(x = i, y = house_num, color = merge_SEQ)) + 
  geom_point() +
  xlab("Record Index") + 
  ylab("House Number") + 
  ggtitle("Sample 2: Merged Sequences (street off)")

seq_st <- ggplot(HN7_st[3100:3400,], aes(x = i, y = house_num, color = merge_SEQ)) + 
  geom_point() +
  xlab("Record Index") + 
  ylab("House Number") + 
  ggtitle("Sample 2: Merged Sequences (street on)")

grid.arrange(sub_seq, seq, seq_st,  ncol = 3)
```


There are more subsequences when `check_street` is on (see `on_` vs `off_`). Some subsequences are merged into ones (see `_merge_SEQ` vs `_SEQ`). `i` is the total number of records.


```{r}
HN7_hn %>% select(i, SEQ, merge_SEQ) %>% full_join(HN7_st %>% select(i, SEQ, merge_SEQ), by = "i") %>%
  rename(off_merge_SEQ = merge_SEQ.x, on_merge_SEQ = merge_SEQ.y,
         off_SEQ = SEQ.x, on_SEQ = SEQ.y) %>%
  select(-i) %>%
  gather(key = type, value = value) %>% unique() %>%
  group_by(type) %>% summarize(cnt = n()) %>% kable()
```

### Sequence number progress

* The top line is merged sequence numbers when street names are also considered; the bottom line is when they are not. 
* Different colors represent different sequnce numbers
* X axis is row numbers

```{r}
#' Column names are changed so that it works with lookAroundBP()
join_df <- HN7_hn %>% select(microfilm, ED, house_num, street_add, merge_SEQ, i) %>%
  full_join(HN7_st %>% select(merge_SEQ, i), by = "i") %>%
  rename(SEQ = merge_SEQ.x,
         merge_SEQ = merge_SEQ.y)

trace <- join_df %>% rename(merge_SEQ_off = SEQ,
                            merge_SEQ_on = merge_SEQ) %>%
  select(i, merge_SEQ_off, merge_SEQ_on) %>% 
  gather(key = seq_type, value = seq_num, -i) %>%
  mutate(dummy_y = ifelse(seq_type=="merge_SEQ_off", 1, 2)) %>%
  filter(!is.na(seq_num))

trace_plot <- ggplot(trace, aes(x=i, y = dummy_y, group = seq_type, color = seq_num)) + 
  geom_line(size = 2) + 
  theme(legend.position = "none") + 
  ylim(c(0, 3))
```


<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }
  .superbigimage img{
     max-width: none;
  }
</style>


#### Row number 1 - 1000
<div class="superbigimage">
```{r , echo = FALSE, fig.width=20,fig.height=3}
trace_plot + xlim(c(1,1000))
```
</div>


```{r}
#' Number of merge sequences
data.frame("street_off" = HN7_hn$merge_SEQ %>% unique() %>% na.omit %>% length(),
           "street_on" = HN7_st$merge_SEQ %>% unique() %>% na.omit %>% length()) %>% kable()
```


## Sequences with and without check_street

There are some records that start new sequences when `check_street` is TRUE, but do not when `check_street` is FALSE (and vice versa). There are 592 such records (out of ~19k). They point to 

* Unstandardized street names e.g.
    * ROOSEVELT ST vs ROOSEVELT STREET (i = 355)
* Trivial mispelling
    * CARL ST vs PEARL ST (i=103)
    * JAMES vs NAMES vs SAMES
    * OLIVER vs OLIWER (i = 1482)
    * FRANKLIN ST vs HUIKAHI ST (i = 4217) **??**
* mistranscribing
    * look around i = 61
    * look around i = 479
    * look around i = 10648
    * Dates as street names e.g. `ADP 18`(i = 410) and `APRL 22`
* Missing values that can be confidently inferred
    * lookaround i = 1198 (ROAD A -- missing -- ROAD B)
* No mistakes

```{r, eval = FALSE}
## table too large
join_df %>% rename(merge_SEQ_off = SEQ,
                        merge_SEQ_on = merge_SEQ) %>%
  select(microfilm, ED, house_num, street_add, 
         i, merge_SEQ_off, merge_SEQ_on) %>% 
  kable %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px")
```


```{r, eval = FALSE}

x <- breakPointI(join_df)
lookAround(join_df, 8280, window_size = 20) %>% View
# LAB_list_2 <- lapply(x, function(pt){
#   return(lookAround(join_df, pt, window_size = 20) %>%
#            rename(merge_SEQ_off = SEQ,
#                   merge_SEQ_on = merge_SEQ))
# })
# View(LAB_list_2[[300]])
# View(LAB_list_2[[301]])
# View(LAB_list_2[[302]])
# View(LAB_list_2[[304]])
# View(LAB_list_2[[307]])
# View(LAB_list_2[[309]])
```


```{r, eval = FALSE, echo = FALSE}
#### 1. `SEQ` and `merge_SEQ` do not start new sequences at the same record

Becuase conditions for the merge are met. 

- No more than two house numbers with a parity different from the rest. Could imply mistranscription
- Records are still in neighboring areas (decreasing and increasing house number in the same area)


LAB_list <- lookAroundBP(HN7_hn, window_size = 6)
do.call(rbind, LAB_list) %>% 
  select(ED, microfilm, house_num, street_add, SEQ, merge_SEQ, seq_par, i) %>% 
  kable() %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px")
```

```{r, eval = FALSE}
selectInput(inputId = "BP", label = 'Choose break point to inspect',
            choices = seq(1,length(x)))
message(input$BP)
renderUI( {
  HTML(LAB_list_2[[input$BP]] %>% kable() %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px"))
})

# renderPlot({
#   par(mar = c(4, 4, .1, .5))
#   hist(
#     faithful$eruptions, as.numeric(input$breaks),
#     col = 'gray', border = 'white',
#     xlab = 'Duration (minutes)', main = ''
#   )
# })
```


```{r, eval = FALSE}
LAB_list_2 <- HN7_hn %>% select(-SEQ) %>% full_join(HN7_st %>% select(merge_SEQ, i), by = "i") %>%
  rename(SEQ = merge_SEQ.x,
         merge_SEQ = merge_SEQ.y) %>% 
  lookAroundBP(.,window_size = 10)
do.call(rbind, LAB_list_2) %>% 
  select(ED, microfilm, house_num, street_add, SEQ, merge_SEQ, seq_par, i) %>% 
  rename(merge_SEQ_off = SEQ,
         merge_SEQ_on = merge_SEQ) %>% 
  kable() %>% kable_styling() %>% 
  kableExtra::scroll_box(width = "100%", height = "400px")

```














.