---
title: "Sequence analysis"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

Using v2 of output.

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(kableExtra)
library(extrafont)

opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

theme_set(
  theme_minimal() +
  theme(text = element_text(family = "Open Sans Light"))
)
```

```{r}
sample <- readRDS("../Data/100k_post_clean.RDS")

source("05_00_getSequence.R")
source("05_00_mergeSequence.R")
```

```{r}
# apply sequence functions

seq_hn <- getMergedSeq(sample, get_check_street = FALSE, merge_check_street = F, merge_check_par = F)
```

## EDA of sequence 'breakpoints'
It might be interesting to look at which house number sequences could have gotten merged if not for a single house number breaking the sequence. These 'breakpoints' could indicate errant house numbers.

```{r}
seq_hn_grp <- seq_hn %>%
  group_by(merge_SEQ) %>%
  summarize(count = n_distinct(hn_1), start = first(hn_1), end = last(hn_1), min = min(hn_1, na.rm = T), max = max(hn_1, na.rm = T)) %>%
  arrange(as.numeric(merge_SEQ))

above <- c(NA, flatten(seq_hn_grp[1:nrow(seq_hn_grp)-1, "end"])) %>%
  unlist()

below <- c(flatten(seq_hn_grp[2:nrow(seq_hn_grp), "start"]), NA) %>%
  unlist()

seq_hn_grp$above_end <- above
seq_hn_grp$below_start <- below
```

Current criteria to check for these 'breakpoints':
1. The sequence only contains one house number  
2. The ending of the sequence above and start of the sequence below have a difference of less than 10   

Data here does not take into account street names
```{r}
filter(seq_hn_grp,
       count == 1
       & abs(above_end - below_start) <= 10) %>%
  mutate(hn = min) %>%
  select(merge_SEQ, hn, above_end, below_start) %>%
#  left_join(select(seq_hn, merge_SEQ, index), by = c("merge_SEQ" = "merge_SEQ")) %>%
#  bind_cols(row = 1: nrow(.)) %>%
  kable() %>% kable_styling() %>%
  scroll_box(height = "300px")

count_this <- filter(seq_hn_grp,
       count == 1
       & abs(above_end - below_start) <= 10) %>%
  mutate(hn = min) %>%
  select(merge_SEQ, hn, above_end, below_start) %>%
  nrow(.)
```

**COUNT: `r count_this`**

### What should these numbers be changed to?

```{r}
plot_seq = function(df, start, end, adjust = c(0, 5)) {
  seq = df %>%
    filter(merge_SEQ %in% seq(start, end))
  
  ggplot(seq) +
    geom_point(aes(x = index, y = hn_1, color = as.factor(merge_SEQ))) +
    geom_text(data = distinct(seq, merge_SEQ, .keep_all = T), 
              aes(x = index + adjust[1], y = hn_1 + adjust[2], 
                  label = paste0(str_to_title(best_match), " (", result_type, ")\n", hn_1)),
              family = "Open Sans Light")
}
```


```{r}
plot_seq(seq_hn, 60, 63, c(1, 1))
```

```{r}
plot_seq(seq_hn, 430, 434, c(0, 10))
```

```{r}
plot_seq(seq_hn, 239, 243, c(0, 10))
```

## Can we confidently change these?

Above and below are within 2
```{r}
filter(seq_hn_grp,
       count == 1
       & abs(above_end - below_start) <= 2) %>%
  mutate(hn = min) %>%
  select(merge_SEQ, hn, above_end, below_start) %>%
#  left_join(select(seq_hn, merge_SEQ, index), by = c("merge_SEQ" = "merge_SEQ")) %>%
#  bind_cols(row = 1: nrow(.)) %>%
  kable() %>% kable_styling() %>%
#  row_spec(c(2, 3, 6, 7, 10, 12, 15, 16, 22, 23, 25, 27, 29, 30, 32, 34, 35, 37), background = "#ff8080") %>%
  scroll_box(height = "300px")
```


Looking specifically at the house numbers
```{r}
narrow_breaks = filter(seq_hn_grp,
       count == 1
       & abs(above_end - below_start) <= 2) %>%
  mutate(merge_SEQ = as.numeric(as.character(merge_SEQ))) %>%
  pull(merge_SEQ)

narrow_breaks_window = narrow_breaks
seqs = seq_hn %>% 
  mutate(merge_SEQ = as.numeric(as.character(merge_SEQ))) %>%
  pull(merge_SEQ) %>% 
  unique()

for (seq in narrow_breaks) {
  seq_ind = which(seqs == seq)
  narrow_breaks_window <- c(narrow_breaks_window, seqs[seq_ind - 1], seqs[seq_ind + 1])
}

narrow_eda = seq_hn %>%
  mutate(merge_SEQ = as.numeric(as.character(merge_SEQ))) %>%
  filter(merge_SEQ %in% narrow_breaks_window) %>%
  select(ED, street_add, best_match, result_type, hn_1, hn_2, hn_3, merge_SEQ) %>%
  distinct()

narrow_eda %>%
  kable() %>% kable_styling() %>%
  row_spec(which(narrow_eda$merge_SEQ %in% narrow_breaks), background = "#ff8080") %>%
  scroll_box(height = "300px")
```

Definitely yes! These are sudden drop in house numbers amongst two extremely close sequences with the same street address (with strong match confidence).
```{r}
plot_seq(seq_hn, 69, 71)
```

```{r}
plot_seq(seq_hn, 278, 280, c(0,0))
```

Maybe?
Original st name was Haull<br>
```{r}
plot_seq(seq_hn, 300, 303)
```

```{r}
plot_seq(seq_hn, 429, 432)
```

## Testing varying parameters of breakpoints

```{r}
narrow_breaks = filter(seq_hn_grp,
       count == 1
       & abs(above_end - below_start) <= 10) %>%
  mutate(merge_SEQ = as.numeric(as.character(merge_SEQ))) %>%
  pull(merge_SEQ)

narrow_breaks_window = narrow_breaks
seqs = seq_hn %>% 
  mutate(merge_SEQ = as.numeric(as.character(merge_SEQ))) %>%
  pull(merge_SEQ) %>% 
  unique()

for (seq in narrow_breaks) {
  seq_ind = which(seqs == seq)
  narrow_breaks_window <- c(narrow_breaks_window, seqs[seq_ind - 1], seqs[seq_ind + 1])
}

narrow_eda = seq_hn %>%
  mutate(merge_SEQ = as.numeric(as.character(merge_SEQ))) %>%
  filter(merge_SEQ %in% narrow_breaks_window) %>%
  select(ED, street_add, best_match, result_type, hn_1, hn_2, hn_3, merge_SEQ) %>%
  distinct()

narrow_eda %>%
  kable() %>% kable_styling() %>%
  row_spec(which(narrow_eda$merge_SEQ %in% narrow_breaks), background = "#ff8080") %>%
  scroll_box(height = "300px")
```



