---
title: "EDA 1880"
author: "Olivia Huang"
date: "11/17/2020"
output:
  html_document:
    df_print: paged
---

***

## Set up
```{r, results='hide', message=FALSE}
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
library(hunspell)
```

```{r, results='hide', message=FALSE}
sample <- read_csv("census_1880_h_mn.csv")
```

```{r, echo=FALSE}
names(sample) <- c("record", "year", "serial", "dwsize", "township", "pageno", "microseq", "n_fam", "splithid", "splitnum", "mcd", "county", "enumdist", "supdist", "street" )

# names(sample) <- c("hnyc_id", "record", "township", "county", "ED", "person_id",
#                    "dwelling_seq", "dwelling_ser", "dwelling_ser2", "hh_seq", "hh_seq_8",
#                    "hh_ser2", "hh_ser_bef_split", "indiv_seq", "split", 
#                    "line_no", "line_no2", "microfilm",
#                    "n_fam", "n_person_bef_split", "house_num", "street_add")
sample <- extract(sample, "street", c("house_num", "street_add"), "(\\d.)(\\D+)")
```
House numbers and street names are separated from the original address column, into house_num and street_add.

## Dataset
Here is a sample of the 1880 census data.
Please note that at the moment, the variable key names are not completely aligned with the 1910 dataset.They need to be configurated before the full run.
```{r, echo=FALSE}
sample_hn <- filter(sample, house_num != 'NA')
head(sample_hn, 100)
```

## Variables
Here's a demonstration of the possible misspellings in the sample data, which is representative of the whole dataset. According to the results, about 70 misspellings are detected in the 100k sample data. Please note that it is possible that some detected words are correctly spelled, such as "albany". A closer look into the street dictionaries will be needed to further clean the street names.
```{r, message=FALSE}
household <- filter(sample, record == "H")

mispellings <- household$street_add %>%
  str_split(pattern = " ") %>%
  unlist() %>%
  unique()

hunspell(mispellings) %>% unlist() %>% unique()
```

## Fill down options
An exploration of fill down using household persons to join addresses. Same attempt was made for 1910 data. 
```{r}
na_street <- which(is.na(household$street_add))
na_street_1 <- na_street - 1
na_street <- c(na_street, na_street_1) %>% 
  unique() %>% 
  sort()
```

```{r, echo=FALSE, message=FALSE}
# sample_filled <- sample %>%
#   fill("dwelling_ser", .direction = "down") %>%
#   filter(record == "P") %>%
#   mutate(dwelling_ser = str_pad(dwelling_ser, 8, "left", pad = 0)) %>%
#   mutate(flag = dwelling_seq != dwelling_ser)
# 
# sum(sample_filled$flag)
```

## House Numbers
Here is a summary of the house numbers and street names in the sample data.
```{r, echo=FALSE, message=FALSE}
tibble(none = nrow(filter(household, is.na(house_num) & is.na(street_add))),
       only_house = nrow(filter(household, !is.na(house_num) & is.na(street_add))),
       only_street = nrow(filter(household, is.na(house_num) & !is.na(street_add))),
       both = nrow(filter(household, !is.na(house_num) & !is.na(street_add)))) %>%
  kable() %>% kable_styling(full_width = FALSE)
```
There are 22843 entries with empty addresses (missing both house number and street name), which are shown as NA in the dataframe. They need to be filled down before geocoding.

