---
title: "EDA 1880 BK"
author: "Olivia Huang"
date: "12/01/2020"
output:
  html_document:
    df_print: paged
---

## Set up
```{r, results='hide', message=FALSE}
library(tidyverse)
library(tidyr)
library(knitr)
library(kableExtra)
library(hunspell)
```

```{r, results='hide', message=FALSE}
sample <- read_csv("/Users/olivia/Documents/GitHub/hnyc_census/1880/census_1880_h_bk_sample100k.csv")
names(sample) <- c("record", "year", "dwelling_ser", "dwsize", "township", "pageno", "microfilm", "n_fam", "hh_ser_bef_split", "split", "mcd", "county", "ED", "supdist", "street" )
sample <- extract(sample, "street", c("house_num", "street_add"), "(\\d.)(\\D+)")

sample_split <- read.csv("/Users/olivia/Documents/GitHub/hnyc_census/1880/Scripts/census_1880_h_bk_split.csv")
```
House numbers and street names are separated from the original address column, into house_num and street_add.

## Dataset
Here is a sample of the 1880 Brooklyn census data.
Variable names are now aligned with 1910 data.
```{r, echo=FALSE}
sample_hn <- filter(sample_split, house_num != 'NA')
head(sample_hn, 100)
```

## Variables
Here's a demonstration of the possible misspellings in the 1880 Brooklyn data, which is representative of the whole dataset. According to the results, about 200 misspellings are detected in the data. Please note that it is possible that some detected words are correctly spelled, such as "albany", "york", "MELROSE". A closer look into the street dictionaries will be needed to further clean the street names.
```{r, message=FALSE}
household1 <- filter(sample_split, record == "H")

mispellings <- household1$street_add %>%
  str_split(pattern = " ") %>%
  unlist() %>%
  unique()

hunspell(mispellings) %>% unlist() %>% unique()
```

## Fill down options
An exploration of fill down using household persons to join addresses. Same attempt was made for 1910 data. 
```{r}
na_street <- which(is.na(household1$street_add))
na_street_1 <- na_street - 1
na_street <- c(na_street, na_street_1) %>% 
  unique() %>% 
  sort()
```

## House Numbers
Here is a summary of the house numbers and street names in the sample data.
There are 8418 entries with empty house numbers and 123572 households with complete addresses. Also, some of the house numbers contain or only contain punctuations, which will be cleaned in later processes.

```{r, echo=FALSE, message=FALSE}
tibble(none = nrow(filter(household1, is.na(house_num) & is.na(street_add))),
       only_house = nrow(filter(household1, !is.na(house_num) & is.na(street_add))),
       only_street = nrow(filter(household1, is.na(house_num) & !is.na(street_add))),
       both = nrow(filter(household1, !is.na(house_num) & !is.na(street_add)))) %>%
  kable() %>% kable_styling(full_width = FALSE)
```
